---
title: "soiltemp_validation"
author: "chaney"
date: "2025-11-24"
output: html_document
---

Validating model soil temperature with NEON data and exploring relationship of Soil T with Soil R
```{r setup, include=FALSE}
# Load libraries

library(DBI)
library(dplyr)
library(geosphere)
library(sp)
library(ggplot2)
library(sf)
library(maps)
library(terra)
library(knitr)
library(tidyr)
library(neonUtilities)
library(swfscMisc)
library(maps)
library(tidyverse)
library(neonSoilFlux)
library(parallel)
library(ncdf4)
library(data.table)
library(purrr)
library(data.table)
library(parallel)
library(lubridate)
library(terra)
library(httr)
library(sp)
library(neonSoilFlux)
library(stringr)
library(scales)
library(stats)


```

---- First, get all of the data -----

Get Soil Temp from SDA Model
```{r}

# Get the coordinates of 6400 sites in Dongchen's output
Neonplots <- read.csv("/usr4/ugrad/chaney/R/3_25/Neon_sites_terrestrial.csv") 

# Add a column for site numbers in Site_Info (site_info has coordinates of each location in df)
Site_Info <- readRDS("/usr4/ugrad/chaney/R/3_25/site.locs.rds")
Site_Info$Site_Number <- seq(1, nrow(Site_Info)) 

# Extract latitude and longitude from Neonplots and Site_Info tables
neon_coords <- Neonplots[ ,2:3] %>%
  select(Latitude = field_latitude, Longitude = field_longitude)
Site_Info_coords <- Site_Info %>%
  select(Site_Number, lat, lon)

# Put Site_Info latitude/longitude in a matrix
Site_Coordinates <- as.matrix(Site_Info[, c("lat", "lon")])

# Same for Neonplots
Neonplots_Coordinates <- as.matrix(Neonplots[, 2:3])

# Make list to store matched list
matched <- data.frame(id = numeric(nrow(Neonplots)), # id in forecast
                      site_id = numeric(nrow(Neonplots))) # NEON id

# Loop through each row in Neonplots and calculate each plot's closest site number
for (i in 1:nrow(Neonplots)) {
  # Extract the coordinates for the current row in Neonplots
  plot_coords <- Neonplots_Coordinates[i, ]
  
  # Compute distances from this plot to all site coordinates (longlat means its in km)
  distances <- spDistsN1(Site_Coordinates, plot_coords, longlat = TRUE)
  
  # Find the index of the minimum distance for closest site
  closest_site_index <- which.min(distances)
  
  # Assign the corresponding Site_Number from Site_Info to the current row in Neonplots
  matched$id[i] <- Site_Info$Site_Number[closest_site_index]
  matched$site_id[i] <- Neonplots$field_site_id[i]
}

# Loop through each row in Neonplots and calculate each plot's closest site number
for (i in 1:nrow(Neonplots)) {
  # Extract the coordinates for the current row in Neonplots
  plot_coords <- Neonplots_Coordinates[i, ]
  
  # Compute distances from this plot to all site coordinates (longlat means its in km)
  distances <- spDistsN1(Site_Coordinates, plot_coords, longlat = TRUE)
  
  # Find the index of the minimum distance for closest site
  closest_site_index <- which.min(distances)
  
  # Assign the corresponding Site_Number from Site_Info to the current row in Neonplots
  matched$id[i] <- Site_Info$Site_Number[closest_site_index]
  matched$site_id[i] <- Neonplots$field_site_id[i]
}

process_ensemble_st <- function(ens) {
  ens_id <- sprintf("%05d", ens)
  st_ens <- list()
  c = 1
  
   # Create a data frame
  st_df <- data.table(
              Time = character(13*365*nrow(matched)),
              ST = numeric(13*365*nrow(matched)),
              Site_ID = character(13*365*nrow(matched))
            )

  # Go through all the sites
  for (i in seq_len(nrow(matched))) {
    
    
    site_id <- matched$site_id[i]
    ID <- matched$id[i]
            
    # Get file path 
 file_path <- paste0(
        "/projectnb/dietzelab/dongchen/anchorSites/NA_runs/SDA_8k_site/batch_Nov_24/Job_", ceiling(ID/267),"/out/ENS-",
        ens_id, "-", ID, "/","/README.txt"
      )
      
       if (file.exists(file_path)== TRUE){
         # Read the text file in 
          temp <- readLines(file_path)
          
         # Get the met data path
          met_data_path <- substr(temp[11],17,166)
          
          # in batch_nogedi not batch
          met_data_path <- str_replace(met_data_path,"batch","batch_nogedi")
          
          # Open the new file
          if (file.exists(met_data_path)== TRUE){
            temp_table <- fread(met_data_path)
            temp_table_yr <- temp_table$V2
            
            for (year in 2012:2024){
              temp_table_yr <- filter(temp_table, V2 == year)

              org_date = paste0(year - 1, "-12-31")
              for (day in 1:366){
                temp_table_yr_day <- filter(temp_table, V3 == day)
                st_df$Site_ID[c] = site_id
                st_df$ST[c] = mean(temp_table_yr_day$V7, na.rm = TRUE)
                
                temp_table[V2 == year & V3 == day, mean(V7, na.rm = TRUE)]
                
                date <- as.Date(day, origin = org_date) 
                st_df$Time[c] <- format(date, "%Y-%m-%d")
                c = c+1
              }
            }
          }
       }
      }
  return(list(st = st_df))
}

Nensem <- 1:100
num_cores <- detectCores() - 1

results_temp <- mclapply(Nensem, process_ensemble_st, mc.cores = num_cores)

st_output <- rbindlist(lapply(results_temp, `[[`, "st"))

# Calculate ensemble mean
st_mean <- st_output[, .(st = mean(st, na.rm = TRUE)), by = .(Time, Site_ID)]

# Merge SR mean value with a time resolution of 3h (Not 1 day)
daily_st_data <- as.data.table(st_mean)

# Format Time to character for processing
daily_st_data[, Time := as.character(Time)]


```

Get NEON Soil Temperature from John's Data in SCC -- Get both column average and surface layer
```{r}

soil_temp <- function(site_number){
  
  years = 2012:2025
  site = matched$site_id[site_number]
  
  # Create list to store data
  all_days <- list() 
  count <- 1 


  for (year in years){
    for (i in 01:12){
      if (i < 10){
        file_name <- paste0("/projectnb/dietzelab/jzobitz/02-NEON-sites/env-data/env-meas-", site, "-",year,"-0",i,".Rda")
      }else{
        file_name <- paste0("/projectnb/dietzelab/jzobitz/02-NEON-sites/env-data/env-meas-", site, "-",year,"-",i,".Rda")
      }
      if (file.exists(file_name)== TRUE){
        load(file_name)
      
        # Use top layer for surface temp
        site_data_surface_temp <- filter(site_data[[2]][[3]], verticalPosition == "501") # In C
        # Use all layers for column temp
        site_data_column_temp <- filter(site_data[[2]][[3]]) # In C
       
        
        # Sort by date
        site_data_surface_temp <- site_data_surface_temp[order(site_data_surface_temp$startDateTime),]
        site_data_column_temp <- site_data_column_temp[order(site_data_column_temp$startDateTime),]
        
        
        # Fine the number of data points per day (48 data points at multiple data sitesand depths for each site)
        dppd_surface = 48 * length(unique(site_data_surface_temp$horizontalPosition))
        
        # Combine positions into one column to find the number of unique ones
        site_data_column_temp$Position <- paste(site_data_column_temp$horizontalPosition,site_data_column_temp$verticalPosition)
        dppd_column = 48 * length(unique(site_data_column_temp$Position))
        
        # Find the number of days 
        n_days_s <- nrow(site_data_surface_temp) / (dppd_surface)
        n_days_c <- nrow(site_data_column_temp) / (dppd_column)
        
        # Add day index (ceil groups every dppd rows)
        site_data_surface_temp$day_idx <- ceiling(seq_len(nrow(site_data_surface_temp)) / dppd_surface)
        site_data_column_temp$day_idx <- ceiling(seq_len(nrow(site_data_column_temp)) / dppd_column)
      
       # Count rows per day
        day_sizes_s <- table(site_data_surface_temp$day_idx)
        day_sizes_c <- table(site_data_column_temp$day_idx)
        
        # Keep only complete days
        complete_days_surface <- as.numeric(names(day_sizes_s[day_sizes_s == dppd_surface]))
        complete_days_column <- as.numeric(names(day_sizes_c[day_sizes_c == dppd_column]))
        
        # Find common days
        common_days <- intersect(complete_days_surface, complete_days_column)
        
        # Filter 
        surface_filt <- site_data_surface_temp %>% filter(day_idx %in% common_days)
        column_filt  <- site_data_column_temp  %>% filter(day_idx %in% common_days)
        
        # Turn into Matrix 
        sts_mat <- matrix(surface_filt$soilTempMean, nrow = dppd_surface)
        stc_mat <- matrix(column_filt$soilTempMean,  nrow = dppd_column)
        
        # Find Times
        surface_filt$startDateTime <- as.POSIXct(surface_filt$startDateTime, tz = "UTC")
        day_times <- tapply(surface_filt$startDateTime, surface_filt$day_idx, min)
        day_times <- day_times[as.character(common_days)]
        day_times <- as.POSIXct(day_times, origin = "1970-01-01", tz = "UTC")
        day_times <- format(as.Date(day_times), "%Y-%m-%d")

        
        # Find daily means ST
        daily_means_sts <- colMeans(sts_mat, na.rm = TRUE)
        daily_means_stc <- colMeans(stc_mat, na.rm = TRUE)
        
        # Add to table
        if (length(day_times) == 0) {
            warning("No complete days available for site ", site)
            return(data.frame(Site_ID = site, Time = as.POSIXct(NA),
                              Neon_ST_column = NA, Neon_ST_surface = NA))
        }else{        
        summary_table <- data.frame(
          Site_ID = site,
          Time = day_times,
          Neon_ST_column = daily_means_stc,
          Neon_ST_surface = daily_means_sts
        )}
        
        # Add table to list
      all_days[[count]] <- summary_table
      count <- count + 1
        
        
      }
    }
  }
  # turn list into table
  summary <- rbindlist(all_days)
  
return(list(stm = summary))
}


# Apply parallel calculation for each site
n = 1:nrow(matched)
num_cores <- detectCores() - 1

# Soil_temp Results
results_temp <- mclapply(n, soil_temp, mc.cores = num_cores)

# Combine results from all ensembles
soil_t <- rbindlist(lapply(results_temp, `[[`, "stm"),fill=TRUE)


```

Combine into table
```{r}

# Combine two soil temperature tables
soil_t_table <-  merge(daily_st_data, soil_t, by = c("Time","Site_ID"))

# Change some names
names(soil_t_table)[names(soil_t_table) == "st"] <- "Model_ST"

# Make time a character vector for soil R
soil_t_table$Time <- as.Date(soil_t_table$Time)
daily_sr$Time <- format(daily_sr$Time, format = "%Y-%m-%d")

# Merge with soil respiration 
soil_table <- merge(daily_sr, soil_t_table, by = c("Time","Site_ID"))

```

---- Next, Look at NEON relationship of Soil R as a function of Soil T 

Soil respiration vs soil temperature -- All data together 
(using NEON column data)
```{r}

# Sort into x order 
soil_table <- soil_table[order(soil_table$Neon_ST_column)]


# Make new data frame & remove infs
df_t_ma <- data.frame(x = soil_table$Neon_ST_column,
                   y = soil_table$Marshall_SR)

df_t_ma <- df_t_ma %>%
  filter(if_all(where(is.numeric), is.finite))

df_t_mq <- data.frame(x = soil_table$Neon_ST_column,
                   y = soil_table$Mill_Quirk_SR)

df_t_mq <- df_t_mq %>%
  filter(if_all(where(is.numeric), is.finite))

# Get rid of rows with NAs

# Loess Regressions
loess90_t_ma <- loess(y ~ x, data=df_t_ma, span=.9, na.action = na.omit)
df_t_ma$y_fit <- predict(loess90_t_ma)

loess90_t_mq <- loess(y ~ x, data=df_t_mq, span=.9, na.action = na.omit)
df_t_mq$y_fit <- predict(loess90_t_mq)

# Scatter Plot with Fitted line
ggplot()+
  geom_point(data = df_t_ma, aes(x = x, y = y, color = "Marshall Data"), size = 0.5) +
  geom_point(data = df_t_mq, aes(x = x, y = y, color = "Millington Quirk Data"), size = 0.5) +
  geom_line(data = df_t_ma, aes(x = x, y = y_fit, color = "Marshall Loess Regression"), size = 1) +
  geom_line(data = df_t_mq, aes(x=x, y=y_fit, color ="Millington Quirk Loess Regression"), size = 1)+
  scale_color_manual(values = c("Marshall Data" = "darkgrey", "Millington Quirk Data" = "lightgrey", "Marshall Loess Regression" = "darkgreen", "Millington Quirk Loess Regression" = "lightgreen")) +
  labs(title = "Soil Respiration as a function of Soil Temperature", x = "Soil Temperature (C)", y = "Soil Respiration (kg/m^2*s)", color = "Dataset")

```


Line fits for each Site NEON Soil Temp
```{r}

sites <- unique(soil_table$Site_ID)

for (i in 1:length(sites)){
  temp_data <- filter(soil_table, Site_ID == sites[i])
  
  temp_data <- temp_data[order(temp_data$Neon_ST_column),]
  
  df_ma <- data.frame(x = temp_data$Neon_ST_column,
                   y = temp_data$Marshall_SR)
  df_ma <- df_ma %>%
           filter(if_all(where(is.numeric), is.finite))

  # Scatter Plot with Fitted line
  print(ggplot(df_ma, aes(x = x, y = y)) + geom_point(size = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = sites[i], x = "Soil Temp", y = "Soil Respiration (Marshall)"))
  
  df_mq <- data.frame(x = temp_data$Neon_ST_column,
                   y = temp_data$Mill_Quirk_SR)
  df_mq <- df_mq %>%
            filter(if_all(where(is.numeric), is.finite))

  # Scatter Plot with Fitted line
  print(ggplot(df_mq, aes(x = x, y = y)) + geom_point(size = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = sites[i], x = "Soil Temp", y = "Soil Respiration (Millington-Quirk"))

}
```


Fit Sigmoidal curve to Soil R as a function of Soil T for NEON and model. Just plotting NEON
```{r}
# Making graph of NEON Soil Respiration as a function of time with Loess Regression and Sigmoidal Fit

sites <- unique(soil_table$Site_ID)

# Make table for asymptote data
sig_fit_vals <- data.frame(Site_ID = character(length(sites)),
                                  Asym_ma = numeric(length(sites)),
                                  Asym_mq = numeric(length(sites)),
                                  Asym_model = numeric(length(sites)),
                                  Xmid_ma = numeric(length(sites)),
                                  Xmid_mq = numeric(length(sites)),
                                  Xmid_model = numeric(length(sites)))

# Make a nested data frame for all temp fitting data
  temp_fits <- list()

# Set to NA (will be changed if not NA)
  sig_fit_vals$Asym_ma <- NA_real_
  sig_fit_vals$Asym_mq <- NA_real_
  sig_fit_vals$Asym_model <- NA_real_
  sig_fit_vals$Xmid_ma <- NA_real_
  sig_fit_vals$Xmid_mq <- NA_real_
  sig_fit_vals$Xmid_model <- NA_real_
  
for (i in 1:length(sites)){
  
  sig_fit_vals$Site_ID[i] <- sites[i]
  
  # Filter for site
  temp_data <- filter(soil_table, Site_ID == sites[i])

  # Sort by x
  temp_data <- temp_data[order(temp_data$Neon_ST_column),]
  
  # Get 2 data sets for Marshall and MQ
  temp_data_ma <- data.frame(x = temp_data$Neon_ST_column,
                     y = temp_data$Marshall_SR,
                     Time = temp_data$Time,
                     Site_ID = temp_data$Site_ID)
  temp_data_ma <- temp_data_ma %>%
  filter(if_all(where(is.numeric), is.finite))
  
  temp_data_mq <- data.frame(x = temp_data$Neon_ST_column,
                     y = temp_data$Mill_Quirk_SR,
                     Time = temp_data$Time,
                     Site_ID = temp_data$Site_ID)
  
  temp_data_mq <- temp_data_mq %>%
  filter(if_all(where(is.numeric), is.finite))
  
  # And one for model
  temp_data_model <- data.frame(x = temp_data$Model_ST,
                     y = temp_data$Model_SR,
                     Time = temp_data$Time,
                     Site_ID = temp_data$Site_ID)
  
  temp_data_model <- temp_data_model %>%
  filter(if_all(where(is.numeric), is.finite))
  
  
  # Do loess regression of all
    # NEON
  loess90_temp_ma <- loess(y ~ x, data=temp_data_ma, span=.9, na.action = na.omit)
  temp_data_ma$y_fit_loess <- predict(loess90_temp_ma)

  loess90_temp_mq <- loess(y ~ x, data=temp_data_mq, span=.9,na.action = na.omit)
  temp_data_mq$y_fit_loess <- predict(loess90_temp_mq)
  
    # MODEL
  loess90_temp_model <- loess(y ~ x, data=temp_data_model, span=.9, na.action = na.omit)
  temp_data_model$y_fit_loess <- predict(loess90_temp_model)
  
  # fit a Sigmoidal curve, if it fails then NA will be default value
  
  tryCatch({
  # First, for Marshall SR
  
  # Use SSlogis to automatically set the start poitns
  sig_fit_t_ma <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = temp_data_ma)
  
  temp_data_ma$y_fit_sig <- predict(sig_fit_t_ma)
  
  # Put asymptote value in table
    coefs_ma<- coef(sig_fit_t_ma)
    sig_fit_vals$Asym_ma[i] <- unname(coefs_ma["Asym"])
    sig_fit_vals$Xmid_ma[i] <- unname(coefs_ma["xmid"])
  
    
  }, error = function(e){
     print(paste("NLS fit failed for data set", i, ":", conditionMessage(e)))
  })
  
   # Then, for Mill Quirk
  tryCatch({
  sig_fit_t_mq <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = temp_data_mq)
  
  temp_data_mq$y_fit_sig <- predict(sig_fit_t_mq)
  
   # Put asymptote value in table
    coefs_mq<- coef(sig_fit_t_mq)
    sig_fit_vals$Asym_mq[i] <- unname(coefs_mq["Asym"])
    sig_fit_vals$Xmid_mq[i] <- unname(coefs_mq["xmid"])
  
    
  }, error = function(e){
        print(paste("NLS fit failed for data set", i, ":", conditionMessage(e)))
  })
  
  # Last, for model
  tryCatch({
    
  sig_fit_model <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = temp_data_model)
  
  temp_data_model$y_fit_sig <- predict(sig_fit_model)
  
    # Put asymptote value in table
    coefs_model<- coef(sig_fit_model)
    sig_fit_vals$Asym_model[i] <- unname(coefs_model["Asym"])
    sig_fit_vals$Xmid_model[i] <- unname(coefs_model["xmid"])
    
  }, error = function(e){
        print(paste("NLS fit failed for data set", i, ":", conditionMessage(e)))
  })
  
  # Graph if the fitting worked
  
  if(is.na(sig_fit_vals$Asym_ma[i]) == FALSE && is.na(sig_fit_vals$Asym_mq[i]) == FALSE){
    
      print(ggplot()+
   geom_point(data = temp_data_ma, aes(x = x, y = y, color = "Marshall Data"), size = 0.5) +
   geom_point(data = temp_data_mq, aes(x = x, y = y, color = "Mill Quirk Data"), size = 0.5) +
    geom_line(data = temp_data_ma, aes(x = x, y = y_fit_sig, color = "Marshall Sigmoidal Fit"), size = 1)+
   geom_line(data = temp_data_mq, aes(x = x, y = y_fit_sig, color = "Mill Quirk Sigmoidal Fit"), size = 1)+
     geom_line(data = temp_data_ma, aes(x = x, y = y_fit_loess, color = "Marshall Loess Fit"), size = 1) +
    geom_line(data = temp_data_mq, aes(x = x, y = y_fit_loess, color = "Mill Quirk Loess Fit"), size = 1) +
 scale_color_manual(values = c("Marshall Data" = "darkgrey", "Mill Quirk Data" = "lightgrey", "Marshall Loess Fit" = "darkgreen", "Mill Quirk Loess Fit" = "lightgreen", "Marshall Sigmoidal Fit" = "royalblue", "Mill Quirk Sigmoidal Fit" = "lightblue")) +
    labs(title = sites[i], x = "Soil Temp", y = "Soil Respiration", color = "Dataset"))
    
  }else if(is.na(sig_fit_vals$Asym_ma[i]) == FALSE && is.na(sig_fit_vals$Asym_mq[i]) == TRUE){
    
      print(ggplot()+
   geom_point(data = temp_data_ma, aes(x = x, y = y, color = "Marshall Data"), size = 0.5) +
    geom_line(data = temp_data_ma, aes(x = x, y = y_fit_sig, color = "Marshall Sigmoidal Fit"), size = 1)+
   geom_line(data = temp_data_ma, aes(x = x, y = y_fit_loess, color = "Marshall Loess Fit"), size = 1) +
    geom_line(data = temp_data_mq, aes(x = x, y = y_fit_loess, color = "Mill Quirk Loess Fit"), size = 1) +
   scale_color_manual(values = c("Marshall Data" = "darkgrey", "Mill Quirk Data" = "lightgrey", "Marshall Loess Fit" = "darkgreen", "Mill Quirk Loess Fit" = "lightgreen", "Marshall Sigmoidal Fit" = "royalblue", "Mill Quirk Sigmoidal Fit" = "lightblue")) +
    labs(title = sites[i], x = "Soil Temp", y = "Soil Respiration", color = "Dataset"))
    
  }else  if(is.na(sig_fit_vals$Asym_ma[i]) == TRUE && is.na(sig_fit_vals$Asym_mq[i]) == FALSE){
    
      print(ggplot()+
   geom_point(data = temp_data_mq, aes(x = x, y = y, color = "Mill Quirk Data"), size = 0.5) +
   geom_line(data = temp_data_mq, aes(x = x, y = y_fit_sig, color = "Mill Quirk Sigmoidal Fit"), size = 1)+
   geom_line(data = temp_data_ma, aes(x = x, y = y_fit_loess, color = "Marshall Loess Fit"), size = 1) +
    geom_line(data = temp_data_mq, aes(x = x, y = y_fit_loess, color = "Mill Quirk Loess Fit"), size = 1) +
   scale_color_manual(values = c("Marshall Data" = "darkgrey", "Mill Quirk Data" = "lightgrey", "Marshall Loess Fit" = "darkgreen", "Mill Quirk Loess Fit" = "lightgreen", "Marshall Sigmoidal Fit" = "royalblue", "Mill Quirk Sigmoidal Fit" = "lightblue")) +
    labs(title = sites[i], x = "Soil Temp", y = "Soil Respiration", color = "Dataset"))
    
  }
  
  # Rename x and y 
  colnames(temp_data_ma)[colnames(temp_data_ma) == "y"] <- "Marshall_SR"
  colnames(temp_data_ma)[colnames(temp_data_ma) == "x"] <- "NEON_ST"
  colnames(temp_data_ma)[colnames(temp_data_ma) == "y_fit_loess"] <- "y_fit_loess_ma"
  colnames(temp_data_ma)[colnames(temp_data_ma) == "y_fit_sig"] <- "y_fit_sig_ma"
  
  colnames(temp_data_mq)[colnames(temp_data_mq) == "y"] <- "Mill_Quirk_SR"
  colnames(temp_data_mq)[colnames(temp_data_mq) == "x"] <- "NEON_ST"
  colnames(temp_data_mq)[colnames(temp_data_mq) == "y_fit_loess"] <- "y_fit_loess_mq"
  colnames(temp_data_mq)[colnames(temp_data_mq) == "y_fit_sig"] <- "y_fit_sig_mq"
  
  colnames(temp_data_model)[colnames(temp_data_model) == "y"] <- "Model_SR"
  colnames(temp_data_model)[colnames(temp_data_model) == "x"] <- "Model_ST"
  colnames(temp_data_model)[colnames(temp_data_model) == "y_fit_loess"] <- "y_fit_loess_model"
  colnames(temp_data_model)[colnames(temp_data_model) == "y_fit_sig"] <- "y_fit_sig_model"
  
  
  # Combine ma and mq data frames
temp_df <- merge(temp_data_ma, temp_data_mq, by = c("Time", "Site_ID", "NEON_ST"))
  
  # Combine with model
  temp_df_final <- merge(temp_df, temp_data_model, by = c("Time", "Site_ID"))
  
  temp_fits[[i]] <- temp_df_final
  
}

```

Try to find something to correlate to Asymptotes
```{r}

# Add lat and long site coordinates

sig_fit_vals$Site_Lat <- numeric(nrow(sig_fit_vals))
sig_fit_vals$Site_Long <- numeric(nrow(sig_fit_vals))


for (i in 1:length(sig_fit_vals$Site_ID)){
  site_num = which(Neonplots$field_site_id == sig_fit_vals$Site_ID[i])
  if (length(site_num)>0){
    sig_fit_vals$Site_Lat[i] <- Neonplots$field_latitude[site_num]
    sig_fit_vals$Site_Long[i] <- Neonplots$field_longitude[site_num]
  }
}

dat = terra::rast("all_data_layers.tif")


# -- Soil Tavg
  
sig_fit_vals$Soil_tavg <- numeric(nrow(sig_fit_vals))

for (i in 1:nrow(sig_fit_vals)){
  temp_data <-  filter(soil_table, Site_ID == sig_fit_vals$Site_ID[i])
  sig_fit_vals$Soil_tavg[i] <- mean(temp_data$Neon_ST_column)
}

 # Graph of Soil Tavg and Asymptote values
  
  print(ggplot()+
    geom_point(data = sig_fit_vals, aes(x = Soil_tavg, y = Asym_ma, color = "Marshall Asymptotes"))+
    geom_point(data = sig_fit_vals, aes(x = Soil_tavg, y = Asym_mq, color = "Mill Quirk Asymptotes"))+
   #geom_smooth(data = sig_fit_vals, aes(x = Soil_tavg, y = Asym_ma, color = "red"), method = "lm", se = FALSE)+
   #geom_smooth(data = sig_fit_vals, aes (x = Soil_tavg, y = Asym_mq, color = "orange"), method = "lm", se = FALSE)+
    scale_color_manual(values = c("Marshall Asymptotes" = "darkgreen", "Mill Quirk Asymptotes" = "lightgreen")) +
    labs(title = "Correlation of Average Temperature to Asymptotes of Soil T as a Function of Soil R fits", x = "Soil T Avg", y = "Asymptote of Sigmoidal fit of Soil R as a function of soil T", color = "Dataset"))

# -- 13 factors 
  
sig_fit_vals$ysd <- numeric(nrow(sig_fit_vals))
sig_fit_vals$abg<- numeric(nrow(sig_fit_vals))
sig_fit_vals$fia <- numeric(nrow(sig_fit_vals))
sig_fit_vals$gedi<- numeric(nrow(sig_fit_vals))
sig_fit_vals$twi<-numeric(nrow(sig_fit_vals))
sig_fit_vals$tavg <- numeric(nrow(sig_fit_vals))
sig_fit_vals$srad <- numeric(nrow(sig_fit_vals))
sig_fit_vals$prec <- numeric(nrow(sig_fit_vals))
sig_fit_vals$vapr <- numeric(nrow(sig_fit_vals))
sig_fit_vals$ph <- numeric(nrow(sig_fit_vals))
sig_fit_vals$n <-numeric(nrow(sig_fit_vals))
sig_fit_vals$soc <- numeric(nrow(sig_fit_vals))
sig_fit_vals$sand <- numeric(nrow(sig_fit_vals))

pt = numeric(1)

# Extract data for each from each Neon site coordinate
for (i in 1:nrow(sig_fit_vals)){
  pt = terra::extract(dat,data.frame(lon=sig_fit_vals$Site_Long[i], lat = sig_fit_vals$Site_Lat[i]))
  sig_fit_vals$ysd[i] <- pt$year_since_disturb
  sig_fit_vals$abg[i] <- pt$agb
  sig_fit_vals$fia[i] <- pt$fia
  sig_fit_vals$gedi[i] <- pt$gedi
  sig_fit_vals$twi[i] <- pt$twi
  sig_fit_vals$tavg[i] <- pt$tavg
  sig_fit_vals$srad[i] <- pt$srad
  sig_fit_vals$prec[i] <- pt$prec
  sig_fit_vals$vapr[i] <- pt$vapr
  sig_fit_vals$ph[i] <- pt$PH
  sig_fit_vals$n[i] <- pt$N
  sig_fit_vals$soc[i] <- pt$SOC
  sig_fit_vals$sand[i] <- pt$Sand
}


# For correlation Coefficient
for (i in 6:18){
  # Fit a line
  sig_fit_vals$temp = sig_fit_vals[[i]]
  fit = lm(sig_fit_vals$temp ~ sig_fit_vals$Asym_ma)
  
  # Graph

  print(ggplot(sig_fit_vals, aes(x = temp, y = Asym_ma)) + 
          geom_point() +
          geom_smooth(method = "lm", se = FALSE) + xlab(colnames(sig_fit_vals[(i)])))

  # Add r squared, slope, p value
}

# -- SOC

# Need to find average Soil Carbon for each site

# Add a row to asymptote data table
sig_fit_vals$SOC_neon <- numeric(nrow(sig_fit_vals))

# Add soil carbon average
for (sitename in sites) {
  # Filter into just site
  all_site <- filter(soil_table_years, Site_ID == sitename)
  
  
  # Find row with that sitename
  row <- which(sig_fit_vals$Site_ID == sitename)
  
  # Add average SOC
  sig_fit_vals$SOC_neon[row]<- mean(all_site$SOC_neon)
}

 print(ggplot()+
    geom_point(data = sig_fit_vals, aes(x = SOC_neon, y = Asym_ma, color = "Marshall Asymptotes"))+
    geom_point(data = sig_fit_vals, aes(x = SOC_neon, y = Asym_mq, color = "Mill Quirk Asymptotes"))+
   geom_smooth(data = sig_fit_vals, aes(x = SOC_neon, y = Asym_ma, color = "Marshall Line Fit"), method = "lm", se = FALSE)+
   geom_smooth(data = sig_fit_vals, aes (x = SOC_neon, y = Asym_mq, color = "Mill Quirk Line Fit"), method = "lm", se = FALSE)+
    scale_color_manual(values = c("Marshall Asymptotes" = "black", "Mill Quirk Asymptotes" = "grey", "Marshall Line Fit" = "darkgreen", "Mill Quirk Line Fit" = "lightgreen")) +
    labs(title = "Correlation of Soil Carbon to Asymptotes of Soil T as a Function of Soil R fits", x = "SOC_neon", y = "Asymptote of of Soil R as a function of Soil Temp", color = "Dataset"))

```

Finding if asympotote/[SOC] correlates to any factors
```{r}

# -- LATITUDE

# Add rows for NEON lat and long
sig_fit_vals$Neon_lat = numeric(nrow(sig_fit_vals))
sig_fit_vals$Neon_long = numeric(nrow(sig_fit_vals))

# Add latitude data from Neonplots
for (i in 1:nrow(sig_fit_vals)) {
  # Find row with that sitename
  row <- which(Neonplots$field_site_id == sig_fit_vals$Site_ID[i])
  
  # Add lat & long if file exists
  if (length(row) > 0 ){
    sig_fit_vals$Neon_lat[i] <- Neonplots$field_latitude[row]
    sig_fit_vals$Neon_long[i] <- Neonplots$field_longitude[row]
  }else{
    sig_fit_vals$Neon_lat[i] <- NA
    sig_fit_vals$Neon_long[i] <- NA
  }
}

# Add a row for asymeptote/SOC
sig_fit_vals$Asym_ma_div_soc <- sig_fit_vals$Asym_ma/sig_fit_vals$SOC_neon
sig_fit_vals$Asym_mq_div_soc <- sig_fit_vals$Asym_mq/sig_fit_vals$SOC_neon

print(ggplot()+
    geom_point(data = sig_fit_vals, aes(x = Neon_lat, y = Asym_ma_div_soc, color = "Marshall Asymptotes"))+
    geom_point(data = sig_fit_vals, aes(x = Neon_lat, y = Asym_mq_div_soc, color = "Mill Quirk Asymptotes"))+
    geom_smooth(data = sig_fit_vals, aes(x = Neon_lat, y = Asym_ma_div_soc, color = "Marshall Line Fit"), method = "lm", se = FALSE)+
   geom_smooth(data = sig_fit_vals, aes (x = Neon_lat, y = Asym_mq_div_soc, color = "Mill Quirk Line Fit"), method = "lm", se = FALSE)+
    scale_color_manual(values = c("Marshall Asymptotes" = "black", "Mill Quirk Asymptotes" = "grey", "Marshall Line Fit" = "darkgreen", "Mill Quirk Line Fit" = "lightgreen")) +
    labs(title = "Correlation of Latitude to Asymptote/SOC", x = "Latitude", y = "Asymptote of Soil R as a function of SOC/ SOC", color = "Dataset"))

# -- LONGITUDE

print(ggplot()+
    geom_point(data = sig_fit_vals, aes(x = Neon_long, y = Asym_ma_div_soc, color = "Marshall Asymptotes"))+
    geom_point(data = sig_fit_vals, aes(x = Neon_long, y = Asym_mq_div_soc, color = "Mill Quirk Asymptotes"))+
    geom_smooth(data = sig_fit_vals, aes(x = Neon_long, y = Asym_ma, color = "Marshall Line Fit"), method = "lm", se = FALSE)+
   geom_smooth(data = sig_fit_vals, aes (x = Neon_long, y = Asym_mq, color = "Mill Quirk Line Fit"), method = "lm", se = FALSE)+
    scale_color_manual(values = c("Marshall Asymptotes" = "black", "Mill Quirk Asymptotes" = "grey", "Marshall Line Fit" = "darkgreen", "Mill Quirk Line Fit" = "lightgreen")) +
    labs(title = "Correlation of Longitude to Asymptote/SOC", x = "Longitude", y = "Asymptote of Soil R as a function of Soil Temp/ SOC", color = "Dataset"))


# -- PRECIPITATION

print(ggplot()+
    geom_point(data = sig_fit_vals, aes(x = prec, y = Asym_ma_div_soc, color = "Marshall Asymptotes"))+
    geom_point(data = sig_fit_vals, aes(x = prec, y = Asym_mq_div_soc, color = "Mill Quirk Asymptotes"))+
    geom_smooth(data = sig_fit_vals, aes(x = prec, y = Asym_ma, color = "Marshall Line Fit"), method = "lm", se = FALSE)+
   geom_smooth(data = sig_fit_vals, aes (x = prec, y = Asym_mq, color = "Mill Quirk Line Fit"), method = "lm", se = FALSE)+
    scale_color_manual(values = c("Marshall Asymptotes" = "black", "Mill Quirk Asymptotes" = "grey", "Marshall Line Fit" = "darkgreen", "Mill Quirk Line Fit" = "lightgreen")) +
    labs(title = "Correlation of Precipitation to Asymptote/SOC", x = "Precipitation", y = "Asymptote of Soil R as a function of Soil Temp/ SOC", color = "Dataset"))

# -- 13 Factors 

col_names = colnames(sig_fit_vals)[4:20]

for (col in col_names){
  print(ggplot()+
    geom_point(data = sig_fit_vals, aes(x = .data[[col]], y = Asym_ma_div_soc, color = "Marshall Asymptotes"))+
    geom_point(data = sig_fit_vals, aes(x = .data[[col]], y = Asym_mq_div_soc, color = "Mill Quirk Asymptotes"))+
    geom_smooth(data = sig_fit_vals, aes(x = .data[[col]], y = Asym_ma, color = "Marshall Line Fit"), method = "lm", se = FALSE)+
   geom_smooth(data = sig_fit_vals, aes (x = .data[[col]], y = Asym_mq, color = "Mill Quirk Line Fit"), method = "lm", se = FALSE)+
    scale_color_manual(values = c("Marshall Asymptotes" = "black", "Mill Quirk Asymptotes" = "grey", "Marshall Line Fit" = "darkgreen", "Mill Quirk Line Fit" = "lightgreen")) +
    labs(title = "Correlation of Precipitation to Asymptote/SOC", x = col, y = paste0("Asymptote of Soil R as a function of SOC/", col), color = "Dataset"))
}


```


---- Now, Validating Model ---- 

 -- First, basic error validations of just soil T
 
 Graph Predicted-Observed Plots for each site, for column and surface
```{r}
sites <- unique(soil_table$Site_ID)

soil_temp_line_fits<- data.frame(Site_ID = character(length(matched$site_id)),
                                slope_column = numeric(length(matched$site_id)),
                                intercept_column = numeric(length(matched$site_id)),
                                slope_surface = numeric(length(matched$site_id)),
                                intercept_surface = numeric(length(matched$site_id)))


for (i in 1:length(sites)){
  temp_data <- filter(soil_table, Site_ID == sites[i])
  
  temp_data <- temp_data[order(temp_data$Model_ST)]
  
 # Fit line
  line_fit_column <- lm(Neon_ST_column~Model_ST, data = temp_data, na.action = na.exclude)
  line_fit_surface <- lm(Neon_ST_surface~Model_ST, data = temp_data, na.action = na.exclude)
  
  temp_data$y_fit_surface <- predict(line_fit_surface)
  temp_data$y_fit_column <- predict(line_fit_column)
  temp_data <- temp_data %>% arrange(Model_ST)
  
  print(ggplot() + 
        geom_point(data = temp_data, aes(x = Model_ST, y = Neon_ST_surface, color = "Surface"))+ 
        geom_point(data = temp_data, aes(x = Model_ST, y = Neon_ST_column, color = "Column"))+ 
        geom_line(data = temp_data, aes(x = Model_ST, y = y_fit_surface, color = "Surface Fit", group = 1), size = 1) +
       geom_line(data = temp_data, aes(x = Model_ST, y = y_fit_column, color = "Column Fit",group = 1), size = 1) + 
        geom_abline(slope = 1, intercept = 0, color = "red", size = 2)+
        scale_color_manual(values = c("Surface" = "lightblue", "Column" = "grey", "Surface Fit" = "blue", "Column Fit" = "black")) +
        labs(title = paste0("Soil Temperature at ", sites[i]), x = "Predicted (C)", y = "Observed (C)"))
  
  
  coefs_column<- coef(line_fit_column)
  coefs_surface<- coef(line_fit_surface)
  
  # add to table
  soil_temp_line_fits$Site_ID[i] = sites[i];
  soil_temp_line_fits$slope_column[i] = unname(coefs_column["(Intercept)"]);
  soil_temp_line_fits$intercept_column[i] = unname(coefs_column["Model_ST"]);
  soil_temp_line_fits$slope_surface[i] = unname(coefs_surface["(Intercept)"]);
  soil_temp_line_fits$intercept_surface[i] = unname(coefs_surface["Model_ST"]);

}
```

 Finding error
```{r}

# Calculate error
soil_table$STemp_error_surface <- soil_table$Model_ST - soil_table$Neon_ST_surface
soil_table$STemp_error_column <- soil_table$Model_ST - soil_table$Neon_ST_column

# Keep a permanent soil_table
soil_table_perm <- soil_table

# Make table for summary statistics
summary_stats_T <- data.frame(Site_ID = character(length(sites)),
                           STemp_corr_coef_surface = numeric(length(sites)),
                           STemp_rms_error_surface = numeric(length(sites)),
                           STemp_corr_coef_column = numeric(length(sites)),
                           STemp_rms_error_column = numeric(length(sites)))


sites <- unique(soil_table$Site_ID)

for (i in 1:length(sites)){
  
  temp_data <- filter(soil_table, Site_ID == sites[i])
 
  
  # Remove inf values
  temp_data <- temp_data %>%
  filter(if_all(where(is.numeric), is.finite))
  
  
  # Add site name
  summary_stats_T$Site_ID[i] <- sites[i]
  
  # Find correlation coeffs
  summary_stats_T$STemp_corr_coef_surface[i] <- cor(temp_data$Neon_ST_surface, temp_data$Model_ST)
  summary_stats_T$STemp_corr_coef_column[i] <- cor(temp_data$Neon_ST_column, temp_data$Model_ST)
  
  # Find RMS errors
  summary_stats_T$STemp_rms_error_surface[i] <- sqrt(mean((temp_data$Neon_ST_surface - temp_data$Model_ST)^2))
  summary_stats_T$STemp_rms_error_column[i] <- sqrt(mean((temp_data$Neon_ST_column - temp_data$Model_ST)^2))
}

```

Make map of corr coef
```{r}

# Add site coordinates
summary_stats_T$Site_Lat <- numeric(length(sites))
summary_stats_T$Site_Long <- numeric(length(sites))

for (i in 1:length(summary_stats_T$Site_ID)){
  site_num = which(Neonplots$field_site_id == summary_stats_T$Site_ID[i])
  summary_stats_T$Site_Lat[i] <- Neonplots$field_latitude[site_num]
  summary_stats_T$Site_Long[i] <- Neonplots$field_longitude[site_num]
}


USA = map_data("world") %>% filter(region == "USA")
us_map <- st_as_sf(USA, 
            coords = c("long", "lat"), 
            crs = 4326)  # CRS 4326 is WGS84 (latitude/longitude)

# Map of correlation coefficient
print(ggplot()+
  geom_polygon(data = USA, aes(x= long, y = lat, group = group), fill="forestgreen", alpha = 0.3) + 
  geom_point(data = summary_stats_T, aes(x= Site_Long, y = Site_Lat, size = STemp_corr_coef_surface, color = STemp_corr_coef_surface)) +
  geom_point(alpha = 0.7) + 
    xlim(-170, -50) + 
    scale_size_continuous(range = c(1,10)) + 
    theme_void() + labs(title="Soil Temperature (Surface) Correlation Coefficient") + 
    theme(plot.title = element_text(hjust = 0.5),))

print(ggplot()+
  geom_polygon(data = USA, aes(x= long, y = lat, group = group), fill="forestgreen", alpha = 0.3) + 
    geom_point(data = summary_stats_T, aes(x= Site_Long, y = Site_Lat, size = STemp_corr_coef_column, color = STemp_corr_coef_column)) +
    geom_point(alpha = 0.7) + 
    xlim(-170, -50) + 
    scale_size_continuous(range = c(1,10)) + 
    theme_void() + labs(title="Soil Temperature (Column) Correlation Coefficient") + 
    theme(plot.title = element_text(hjust = 0.5),))

```

Time series
```{r}
for (i in 1:nrow(matched)){
  # Make a graph of a year of one of them
  temp_data <- filter(soil_table, Site_ID == matched$site_id[i])
  
  print(ggplot(temp_data) +  
    geom_line(aes(x = Time, y = Model_ST,group = 1), linewidth = 1, color = "orange") +
    geom_line(aes(x = Time, y = Neon_ST_surface, group = 1), linewidth = 1, color = "darkgreen") + 
    labs(title = matched$site_id[i], x = "date", y = "Soil Temperature"))
}
```


-- Next, validating Soil R as a function of Soil T relationship 
 
Loess Regressions to compare NEON and model soil temperatures -- for surface and column temps -- SITES TOGETHER
```{r}

# Loess Regressions for Model 

df_t_model <- data.frame(x = soil_table$Model_ST,
                   y = soil_table$Model_SR)

loess90_t_model <- loess(y ~ x, data=df_t_model, span=.9)
df_t_model$y_fit <- predict(loess90_t_model)

ggplot() +
  geom_line(data = df_t_ma, aes(x = x, y = y_fit, color = "NEON Data: Marshall Method"), size = 1) +
  geom_line(data = df_t_mq, aes(x = x, y = y_fit, color = "NEON Data: Mill Quirk Method"), size = 1) +
  geom_line(data = df_t_model, aes(x = x, y = y_fit, color = "Model Output"), size = 1)+
  scale_color_manual(values = c("Model Output" = "orange", "NEON Data: Mill Quirk Method" = "lightgreen", "NEON Data: Marshall Method" = "darkgreen")) +
  labs(title = "Regression Model for Soil R as a function of T",
       x = "Soil Temperature",
       y = "Soil Respiration",
       color = "Dataset")

```
 
 Line fits for each Model Temp
```{r}

for (i in 1:length(sites)){
  temp_data <- filter(soil_table, Site_ID == sites[i])
  
  temp_data <- temp_data[order(temp_data$Model_ST)]
  
  df <- data.frame(x = temp_data$Model_ST,
                   y = temp_data$Model_SR)

  # Scatter Plot with Fitted line
  print(ggplot(df, aes(x = x, y = y)) + geom_point(size = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = sites[i], x = "Soil Temp", y = "Soil Respiration"))
}

```
 

Perform LOESS Regression for NEON & Model at each site temperature
```{r}
for (i in 1:length(sites)){
  temp_data_ma <- filter(soil_table, Site_ID == sites[i])
  temp_data_mq <- filter(soil_table, Site_ID == sites[i])
  temp_data_model <- filter(soil_table, Site_ID == sites[i])
  
  temp_data_ma <- temp_data_ma[order(temp_data_ma$Neon_ST_column)]
  temp_data_mq <- temp_data_mq[order(temp_data_mq$Neon_ST_column)]
  temp_data_model <- temp_data_model[order(temp_data_model$Neon_ST_column)]
  
  df_temp_ma <- data.frame(x = temp_data_ma$Neon_ST_column,
                   y = temp_data_ma$Marshall_SR)
   df_temp_ma  <-   df_temp_ma  %>%
  filter(if_all(where(is.numeric), is.finite))
  
  
  df_temp_mq <- data.frame(x = temp_data_mq$Neon_ST_column,
                   y = temp_data_mq$Mill_Quirk_SR)
    df_temp_mq  <-   df_temp_mq  %>%
  filter(if_all(where(is.numeric), is.finite))
  
  df_temp_model <- data.frame(x = temp_data_model$Model_ST,
                   y = temp_data_model$Model_SR)
    df_temp_model  <-   df_temp_model  %>%
  filter(if_all(where(is.numeric), is.finite))

  # NEON
  loess90_temp_ma <- loess(y ~ x, data=df_temp_ma, span=.9)
  df_temp_ma$y_fit <- predict(loess90_temp_ma)

  loess90_temp_mq <- loess(y ~ x, data=df_temp_mq, span=.9)
  df_temp_mq$y_fit <- predict(loess90_temp_mq)
  
  # SDA Model
  loess90_temp_model <- loess(y ~ x, data=df_temp_model, span=.9)
  df_temp_model$y_fit <- predict(loess90_temp_model)

  # TEMP
  
  print(ggplot() +
    geom_line(data = df_temp_ma, aes(x = x, y = y_fit, color = "NEON Data: Marshall Method"), size = 1) +
    geom_line(data = df_temp_mq, aes(x = x, y = y_fit, color = "NEON Data: Mill Quirk Method"), size = 1) +
    geom_line(data = df_temp_model, aes(x = x, y = y_fit, color = "Model Output"), size = 1)+
    scale_color_manual(values = c("Model Output" = "orange", "NEON Data: Mill Quirk Method" = "lightgreen", "NEON Data: Marshall Method" = "darkgreen")) +
    labs(title = sites[i],
         x = "Soil Temperature",
         y = "Soil Respiration",
         color = "Dataset"))

}
```

Soil respiration vs soil temperature for model & NEON, Each Site with sigmoidal and loess regression fits -- With no reference points
```{r}
for (i in 1:length(sites)){
  
  # Get data from temperature fits 
  df <- temp_fits[[i]]
  
  # Graph model and NEON loess fits
  pp <- ggplot() + 
    geom_line(data = df, aes(x = Model_ST, y = y_fit_loess_model, color = "Model Loess Fit"), size = 1) +
    geom_line(data = df, aes(x = NEON_ST, y = y_fit_loess_ma, color = "Marshall Loess Fit"), size = 1) +
    geom_line(data = df, aes(x = NEON_ST, y = y_fit_loess_mq, color = "Millington Quirk Loess Fit"), size = 1) 
    
  
  # If each sigmoidal fit exists, plot it
  if(is.na(sig_fit_vals$Asym_model[i]) == FALSE){
    pp <- pp +  geom_line(data = df, aes(x = Model_ST, y = y_fit_sig_model, color = "Model Sigmoidal Fit"), size = 1)
  }
  
  if(is.na(sig_fit_vals$Asym_ma[i]) == FALSE){
    pp <- pp +  geom_line(data = df, aes(x = NEON_ST, y = y_fit_sig_ma, color = "Marshall Sigmoidal Fit"), size = 1) 
  }
  
  if(is.na(sig_fit_vals$Asym_mq[i]) == FALSE){
    pp <- pp +  geom_line(data = df, aes(x = NEON_ST, y = y_fit_sig_mq, color = "Millington Quirk Sigmoidal Fit"), size = 1) 
  }
  
  pp <- pp +  scale_color_manual(values = c("Model Loess Fit" = "orange", "Model Sigmoidal Fit" = "orangered", "Marshall Loess Fit" = "darkgreen", "Marshall Sigmoidal Fit" = "darkblue", "Millington Quirk Loess Fit" = "lightgreen", "Millington Quirk Sigmoidal Fit" = "lightblue")) +
    labs(title = sites[i], x = "Soil Temp", y = "Soil Respiration", color = "Dataset")

  # Print the plot 
  print (pp)
    
}
```

Soil respiration vs soil temperature, Normalized sigmoidal fits 
 Reference points at t = 10
```{r}

# Make a table to keep track of how far off each is from NEON values on average
Bias_table <- data.frame(Site_ID = character(length(sites)),
                              Bias_to_Marshall_loess= numeric(length(sites)),
                              Bias_to_Mill_Quirk_loess = numeric(length(sites)),
                              Bias_to_Marshall_sig= numeric(length(sites)),
                              Bias_to_Mill_Quirk_sig = numeric(length(sites)))


for (i in 1:length(sites)){

  temp_data <- temp_fits[[i]]
  
  # -- Normalize the curve
  if ("y_fit_sig_ma" %in% colnames(temp_data) == TRUE){
    max_val_ma <- max(temp_data$y_fit_sig_ma)
    temp_data$y_fit_sig_ma_norm <- temp_data$y_fit_sig_ma/max_val_ma}
  if ("y_fit_sig_mq" %in% colnames(temp_data) == TRUE){
    max_val_mq <- max(temp_data$y_fit_sig_mq)
    temp_data$y_fit_sig_mq_norm <- temp_data$y_fit_sig_mq/max_val_mq}
  if ("y_fit_sig_model" %in% colnames(temp_data) == TRUE){
    max_val_model <- max(temp_data$y_fit_sig_model)
    temp_data$y_fit_sig_model_norm <- temp_data$y_fit_sig_model/max_val_model}
  
  # Graph
  pp <- ggplot() 
   #   geom_line(data = temp_data, aes(x = Model_ST, y = y_fit_loess_model_ref, color = "Model Loess Fit"), size = 1) +
   #   geom_line(data = temp_data, aes(x = NEON_ST, y = y_fit_loess_ma_ref, color = "Marshall Loess Fit"), size = 1) +
   #   geom_line(data = temp_data, aes(x = NEON_ST, y = y_fit_loess_mq_ref, color = "Millington Quirk Loess Fit"), size = 1) 
      
    
    # If each sigmoidal fit exists, plot it
    if("y_fit_sig_ma" %in% colnames(temp_data) == TRUE){
      pp <- pp +  geom_line(data = temp_data, aes(x = NEON_ST, y = y_fit_sig_ma_norm, color = "Millington Quirk Sigmoidal Fit"), size = 1)
    }
    
    if("y_fit_sig_mq" %in% colnames(temp_data) == TRUE){
      pp <- pp +  geom_line(data = temp_data, aes(x = NEON_ST, y = y_fit_sig_mq_norm, color = "Marshall Sigmoidal Fit"), size = 1) 
    }
    
    if("y_fit_sig_model" %in% colnames(temp_data) == TRUE){
      pp <- pp +  geom_line(data = temp_data, aes(x = Model_ST, y = y_fit_sig_model_norm, color = "Model Sigmoidal Fit"), size = 1) 
    }
    
    pp <- pp +  scale_color_manual(values = c("Model Loess Fit" = "orange", "Model Sigmoidal Fit" = "orangered", "Marshall Loess Fit" = "darkgreen", "Marshall Sigmoidal Fit" = "darkblue", "Millington Quirk Loess Fit" = "lightgreen", "Millington Quirk Sigmoidal Fit" = "lightblue")) +
      labs(title = sites[i], x = "Soil Temp", y = "Soil Respiration", color = "Dataset")
  
    # Print the plot 
    print (pp)
    
    # Reassign temp fits so that it has reference data
    temp_fits[[i]] <- temp_data
}



```


Soil respiration vs soil temperature, NORMALIZED sigmoidal fits -- With reference points ** NEW METHOD
 Reference points at t = 10
```{r}
# Make a table to keep track of how far off each is from NEON values on average
Bias_table <- data.frame(Site_ID = character(length(sites)),
                              Bias_to_Marshall_loess= numeric(length(sites)),
                              Bias_to_Mill_Quirk_loess = numeric(length(sites)),
                              Bias_to_Marshall_sig= numeric(length(sites)),
                              Bias_to_Mill_Quirk_sig = numeric(length(sites)))


for (i in 1:length(sites)){

  # Get data from temp fits
  temp_data <- temp_fits[[i]]
  
   # set the number
  target_value <- 10 # using 10 becuase it seems present in most data sets
  
       
  # -- Nofrmalize the curve and doing reference
  if ("y_fit_sig_ma" %in% colnames(temp_data) == TRUE){
    max_val_ma <- max(temp_data$y_fit_sig_ma)
    temp_data$y_fit_sig_ma_norm <- temp_data$y_fit_sig_ma/max_val_ma
    
     temp_data_ma <- data.frame(x = temp_data$NEON_ST,
                       y = temp_data$Marshall_SR,
                       Time = temp_data$Time,
                       Site_ID = temp_data$Site_ID)
      temp_data_ma <- temp_data_ma %>%
      filter(if_all(where(is.numeric), is.finite))
    
    sig_fit_t_ma <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = temp_data_ma)
      temp_data_ma$y_fit_sig <- predict(sig_fit_t_ma)
    
    y_ma <- predict(sig_fit_t_ma, data.frame(x = target_value))
    
    temp_data$y_fit_sig_ma_ref <- temp_data$y_fit_sig_ma_norm - y_ma/max_val_ma
    
    }
  if ("y_fit_sig_mq" %in% colnames(temp_data) == TRUE){
    max_val_mq <- max(temp_data$y_fit_sig_mq)
    temp_data$y_fit_sig_mq_norm <- temp_data$y_fit_sig_mq/max_val_mq
    
      temp_data_mq <- data.frame(x = temp_data$NEON_ST,
                         y = temp_data$Mill_Quirk_SR,
                         Time = temp_data$Time,
                         Site_ID = temp_data$Site_ID)
      
      temp_data_mq <- temp_data_mq %>%
      filter(if_all(where(is.numeric), is.finite))
    
     sig_fit_t_mq <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = temp_data_mq)
  temp_data_mq$y_fit_sig <- predict(sig_fit_t_mq)
   
  y_mq <- predict(sig_fit_t_mq, data.frame(x = target_value))
  
  temp_data$y_fit_sig_mq_ref <- temp_data$y_fit_sig_mq_norm - y_mq/max_val_mq
    }
  
  if ("y_fit_sig_model" %in% colnames(temp_data) == TRUE){
    max_val_model <- max(temp_data$y_fit_sig_model)
    temp_data$y_fit_sig_model_norm <- temp_data$y_fit_sig_model/max_val_model
    
     # And one for model
      temp_data_model <- data.frame(x = temp_data$Model_ST,
                         y = temp_data$Model_SR,
                         Time = temp_data$Time,
                         Site_ID = temp_data$Site_ID)
      
      temp_data_model <- temp_data_model %>%
      filter(if_all(where(is.numeric), is.finite))
    
    sig_fit_model <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = temp_data_model)
  temp_data_model$y_fit_sig <- predict(sig_fit_model)
  
   y_model <- predict(sig_fit_model, data.frame(x = target_value))
   
   temp_data$y_fit_sig_model_ref <- temp_data$y_fit_sig_model_norm - y_model/max_val_model
    }

  
  # Fill in bias table
  Bias_table$Site_ID[i] <- sites[i]
 
  
  # Graph
  pp <- ggplot() 
   #   geom_line(data = temp_data, aes(x = Model_ST, y = y_fit_loess_model_ref, color = "Model Loess Fit"), size = 1) +
   #   geom_line(data = temp_data, aes(x = NEON_ST, y = y_fit_loess_ma_ref, color = "Marshall Loess Fit"), size = 1) +
   #   geom_line(data = temp_data, aes(x = NEON_ST, y = y_fit_loess_mq_ref, color = "Millington Quirk Loess Fit"), size = 1) 
      
    
    # If each sigmoidal fit exists, plot it
    if("y_fit_sig_ma" %in% colnames(temp_data) == TRUE){
      pp <- pp +  geom_line(data = temp_data, aes(x = NEON_ST, y = y_fit_sig_ma_ref, color = "Millington Quirk Sigmoidal Fit"), size = 1)
    }
    
    if("y_fit_sig_mq" %in% colnames(temp_data) == TRUE){
      pp <- pp +  geom_line(data = temp_data, aes(x = NEON_ST, y = y_fit_sig_mq_ref, color = "Marshall Sigmoidal Fit"), size = 1) 
    }
    
    if("y_fit_sig_model" %in% colnames(temp_data) == TRUE){
      pp <- pp +  geom_line(data = temp_data, aes(x = Model_ST, y = y_fit_sig_model_ref, color = "Model Sigmoidal Fit"), size = 1) 
    }
    
    pp <- pp +  scale_color_manual(values = c("Model Loess Fit" = "orange", "Model Sigmoidal Fit" = "orangered", "Marshall Loess Fit" = "darkgreen", "Marshall Sigmoidal Fit" = "darkblue", "Millington Quirk Loess Fit" = "lightgreen", "Millington Quirk Sigmoidal Fit" = "lightblue")) +
      labs(title = sites[i], x = "Soil Temp", y = "Soil Respiration", color = "Dataset")
  
    # Print the plot 
    print (pp)
    
    # Reassign temp fits so that it has reference data
    temp_fits[[i]] <- temp_data
}



```

Scoring using MSD & Correlation Coefficient -- where overlapping when set to reference point at 10
```{r}
# Make a table for score

sig_fit_scores <- data.frame(Site_ID = character(length(sites)),
                              MSD_ma = numeric(length(sites)),
                              MSD_mq = numeric(length(sites)),
                              CC_ma = numeric(length(sites)),
                              CC_mq = numeric(length(sites)),
                              Bias_ma = numeric (length(sites)),
                              Bias_mq = numeric (length(sites)),
                              Input_overlap = numeric (length(sites)))

# Loop through each site
for (i in 1:length(sites)){

  temp_data <- temp_fits[[i]]
  sig_fit_scores$Site_ID[i] <- sites[i]
  
  # Find where range overlaps if both exist
  if ("y_fit_sig_ma" %in% colnames(temp_data) == TRUE && "y_fit_sig_mq" %in% colnames(temp_data) == TRUE && "y_fit_sig_model" %in% colnames(temp_data) == TRUE){
  
  min_x <- max(min(temp_data$NEON_ST), min(temp_data$Model_ST));
  max_x <- min(max(temp_data$NEON_ST), max(temp_data$Model_ST));
  
    # Find input percent overlap
  percent_overlap <- ((max_x - min_x)/(max(max(temp_data$NEON_ST), max(temp_data$Model_ST))-min(min(temp_data$NEON_ST), min(temp_data$Model_ST)))) * 100
  
  sig_fit_scores$Input_overlap[i] <- round(percent_overlap, digits = 2)

    if ((min_x < max_x) == TRUE){
      
        # Choose 50 points distributed across the range
       overlap <- data.frame(x = seq(from = min_x, to = max_x, length.out = 50))
       
       # Find the values at those points
       # Need to find function of fit again (because these points do not exist in data and don't want to use closest point)
       
       # Make dfs
       temp_data_ma <- data.frame(x = temp_data$NEON_ST,
                       y = temp_data$Marshall_SR,
                       Time = temp_data$Time,
                       Site_ID = temp_data$Site_ID)
      temp_data_ma <- temp_data_ma %>%
      filter(if_all(where(is.numeric), is.finite))
      
      temp_data_mq <- data.frame(x = temp_data$NEON_ST,
                         y = temp_data$Mill_Quirk_SR,
                         Time = temp_data$Time,
                         Site_ID = temp_data$Site_ID)
      
      temp_data_mq <- temp_data_mq %>%
      filter(if_all(where(is.numeric), is.finite))
      
      # And one for model
      temp_data_model <- data.frame(x = temp_data$Model_ST,
                         y = temp_data$Model_SR,
                         Time = temp_data$Time,
                         Site_ID = temp_data$Site_ID)
      
      temp_data_model <- temp_data_model %>%
      filter(if_all(where(is.numeric), is.finite))
       
      # Sig fits
      sig_fit_t_ma <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = temp_data_ma)
      sig_fit_t_mq <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = temp_data_mq)
      sig_fit_t_model <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = temp_data_model)
       
       # want to do scoring when setting reference point, so find values at reference point
      
       target_value <- 10 # using 10 becuase it seems present in most data sets
  
       ref_diff_ma <- predict(sig_fit_t_ma, data.frame(x = target_value))[1]
       ref_diff_mq <- predict(sig_fit_t_mq, data.frame(x = target_value))[1]
       ref_diff_model <- predict(sig_fit_t_model, data.frame(x = target_value))[1]
       
      # Add bias to table
      sig_fit_scores$Bias_ma[i] <- ref_diff_model - ref_diff_ma
      sig_fit_scores$Bias_mq[i] <- ref_diff_model - ref_diff_mq
       
       # find max values to normalize
       max_val_ma <- max(temp_data$y_fit_sig_ma)
       max_val_mq <- max(temp_data$y_fit_sig_mq)
       max_val_model <- max(temp_data$y_fit_sig_model)
       
      # use function to find values at points
       actual <- (predict(sig_fit_t_model, overlap)[1:50] - ref_diff_model)/max_val_ma
       predicted_ma <- (predict(sig_fit_t_ma, overlap)[1:50]- ref_diff_ma)/max_val_mq
       predicted_mq <- (predict(sig_fit_t_mq, overlap)[1:50] - ref_diff_mq)/max_val_model
       
      # Find mean((actual-predicted)^2)
       msd_ma <-  mean((actual-predicted_ma)^2)
       msd_mq <-  mean((actual-predicted_mq)^2)
       sig_fit_scores$MSD_ma[i] <- round(msd_ma, digits = 2)
       sig_fit_scores$MSD_mq[i] <- round(msd_mq, digits = 2)
      
      sig_fit_scores$CC_ma[i] <- round(cor(actual, predicted_ma), digits = 2);
      sig_fit_scores$CC_mq[i] <- round(cor(actual, predicted_mq), digits = 2);
  
  # Put into table to compare
    }else{
      # If range doesn't over lap, add "NA" to table
      sig_fit_scores$MSD_ma[i] <- "No overlap"
      sig_fit_scores$MSD_mq[i] <- "No overlap"
    }

  }else{ # NA = no sig fit
      sig_fit_scores$MSD_ma[i] <- NA
      sig_fit_scores$MSD_mq[i] <- NA
  }
  
}

```


Scoring with iterative closest point


Soil respiration vs soil temperature, Example sites with sigmoidal and loess regression fits -- With reference points
 Reference point at sigmoidal inflection points
```{r}

for (i in 1:length(sites)){

  # Get data for site
  temp_data <- temp_fits[[i]]
  
  # Set up plot
   pp <- ggplot() 

# Find inflection point of sig curve and set as reference

    # Marshall
  if("y_fit_sig_ma" %in% colnames(temp_data) == TRUE){
    
   # Set the x as the inflection point (xmid)
    x_inf_point_ma <- sig_fit_vals$Xmid_ma[i]
    
    
    # Calculate absolute differences between target and temp
    abs_diff_ma <- abs(temp_data$NEON_ST - x_inf_point_ma)
    
    # Find the index of the minimum absolute difference
    closest_row_index_ma <- which.min(abs_diff_ma)
    
    y_inf_point_ma <- temp_data$y_fit_sig_ma[closest_row_index_ma]
    
    # Find new reference points with inflections
    temp_data$x_ma_ref <- temp_data$NEON_ST - x_inf_point_ma
    temp_data$y_fit_sig_ma_ref <- temp_data$y_fit_sig_ma - y_inf_point_ma
    
    # add to plot
     pp <- pp + 
      geom_line(data = temp_data, aes(x = x_ma_ref, y = y_fit_sig_ma_ref, color = "Marshall Sigmoidal Fit"), size = 1)
  }
   
   # Mill Quirk
  if("y_fit_sig_mq" %in% colnames(temp_data) == TRUE){
    
   # Set the x as the inflection point (xmid)
    x_inf_point_mq <- sig_fit_vals$Xmid_mq[i]
    
    # Calculate absolute differences between target and temp
    abs_diff_mq <- abs(temp_data$NEON_ST - x_inf_point_mq)
    
    # Find the index of the minimum absolute difference
    closest_row_index_mq <- which.min(abs_diff_mq)
    
    y_inf_point_mq <- temp_data$y_fit_sig_mq[closest_row_index_mq]
    
    # Find new reference points with inflections
    temp_data$x_mq_ref <- temp_data$NEON_ST - x_inf_point_mq
    temp_data$y_fit_sig_mq_ref <- temp_data$y_fit_sig_mq - y_inf_point_mq
    
    # Add to plot
     pp <- pp + 
      geom_line(data = temp_data, aes(x = x_mq_ref, y = y_fit_sig_mq_ref, color = "Millington Quirk Sigmoidal Fit"), size = 1)
  }
   
    # Model
  if("y_fit_sig_model" %in% colnames(temp_data) == TRUE){
    
   # set the x as the inflection point (xmid)
    x_inf_point_model <- sig_fit_vals$Xmid_model[i]
    
    # Calculate absolute differences between target and temp
    abs_diff_model <- abs(temp_data$Model_ST - x_inf_point_model)
    
    # Find the index of the minimum absolute difference
    closest_row_index_model <- which.min(abs_diff_model)
    
    y_inf_point_model <- temp_data$y_fit_sig_model[closest_row_index_model]
    
    # Find new reference points with inflections
    temp_data$x_model_ref <- temp_data$Model_ST - x_inf_point_model
    temp_data$y_fit_sig_model_ref <- temp_data$y_fit_sig_model - y_inf_point_model
    
    # add to plot
     pp <- pp + 
      geom_line(data = temp_data, aes(x = x_model_ref, y = y_fit_sig_model_ref, color = "Model Sigmoidal Fit"), size = 1)
  }
  
  # If any sigs, plot pp
   
   if (("y_fit_sig_ma" %in% colnames(temp_data) == TRUE)| ("y_fit_sig_mq" %in% colnames(temp_data) == TRUE) | ("y_fit_sig_model" %in% colnames(temp_data) == TRUE)){
     
     pp <- pp +  scale_color_manual(values = c("Model Loess Fit" = "orange", "Model Sigmoidal Fit" = "orangered", "Marshall Loess Fit" = "darkgreen", "Marshall Sigmoidal Fit" = "darkblue", "Millington Quirk Loess Fit" = "lightgreen", "Millington Quirk Sigmoidal Fit" = "lightblue")) +
          labs(title = sites[i], x = "Soil Temp", y = "Soil Respiration", color = "Dataset")
    
    # print the plot
    print(pp)
   }
  
}
  
  
  

```


Chooisng 10 points and finding mean squared distance

Running met2model
```{r}

met2model.SIPNET("/Users/chane/OneDrive/Documents/ef_lab/R_files", )
```



Comparing Neon_ST_column to new calculations from 
```{r}

# load in .clim file that i made

temp_table <- fread("ERA5.1.2012-01-01.2024-01-01.clim")
years = 2012:2024
length = years[length(years)]- years[1] + 2

at_df <- data.table(
              Time = character(length*366),
              Site_ID = character(length*366),
              AT = numeric(length*366),
              ST_og = numeric(length*366),
              ST_20 = numeric(length*366),
              ST_cutoff = numeric(length*366)
            )
c = 1

for (year in years){
        temp_table_yr <- filter(temp_table, V1 == year)
              org_date = paste0(year - 1, "-12-31")
              
        for (day in 1:366){
                temp_table_yr_day <- filter(temp_table, V2 == day)
                at_df$Site_ID[c] = "BART"
                at_df$AT[c] = mean(temp_table_yr_day$V5, na.rm = TRUE)
                at_df$ST_og[c] = mean(temp_table_yr_day$V6, na.rm = TRUE)
                at_df$ST_20[c] = mean(temp_table_yr_day$V7, na.rm = TRUE)
                at_df$ST_cutoff[c] = mean(temp_table_yr_day$V8, na.rm = TRUE)
              
                temp_table[V2 == year & V3 == day, mean(V6, na.rm = TRUE)]
                
                # need to convert date into "year-mo-day format"
                date <- as.Date(day, origin = org_date) 
                at_df$Time[c] <- format(date, "%Y-%m-%d")
                c = c+1
                
         }
}



```


Graph each new attempt
```{r}

  # give soil T a year
soil_t$Year <- substr(soil_t$Time,1,4)
at_df$Year <- substr(at_df$Time,1,4)

# Make into date class
soil_t$Time <- as.Date(soil_t$Time)
at_df$Time <- as.Date(at_df$Time)

  # make a graph of a year of one of them
  temp_data <- filter(soil_t, Site_ID == matched$site_id[1])
  
  
  print(ggplot() +  
    geom_line(data = at_df, aes(x = Time, y = ST_og ,group = 1, color = "Original Model ST"), linewidth = 1) +
    geom_line(data = temp_data, aes(x = Time, y = Neon_ST_column, group = 0.5, color = "NEON ST"), linewidth = 1) + 
    geom_line(data = at_df, aes(x = Time, y = ST_20 ,group = 1, color = "Model ST with 20"), linewidth = 1) +
    geom_line(data = at_df, aes(x = Time, y = ST_cutoff ,group = 1, color = "Model ST with Cutoff"), linewidth = 1) +
    labs(title = matched$site_id[i], x = "date", y = "Soil Temperature") + 
    scale_x_date(limits = as.Date(c("2017-01-01", "2018-12-31"))) +
    scale_color_manual(values = c("Original Model ST" = "orange", "Model ST with 20" = "pink", "Model ST with Cutoff" = "salmon","NEON ST" = "lightgreen")))


```

Graph each new attempt
```{r}

# load in .clim file that i made

temp_table <- fread("ERA5.1.BART.2012-01-01.2024-01-01.clim")
years = 2012:2024
length = years[length(years)]- years[1] + 3

at_df_bart_1 <- data.table(
              Time = character(length*366),
              Site_ID = character(length*366),
              AT = numeric(length*366),
              ST_og = numeric(length*366),
              ST_20 = numeric(length*366),
              ST_25 = numeric(length*366),
              ST_30 = numeric(length*366),
              ST_35 = numeric(length*366),
              ST_40 = numeric(length*366),
              ST_45 = numeric(length*366),
              ST_50 = numeric(length*366),
              ST_55 = numeric(length*366),
              ST_60 = numeric(length*366),
              ST_65 = numeric(length*366),
              ST_70 = numeric(length*366),
              ST_75 = numeric(length*366),
              ST_80 = numeric(length*366),
              ST_85 = numeric(length*366),
              ST_90 = numeric(length*366),
              ST_95 = numeric(length*366),
              ST_100 = numeric(length*366),
              ST_cutoff = numeric(length*366)
            )
c = 1

for (year in years){
        temp_table_yr <- filter(temp_table, V1 == year)
              org_date = paste0(year - 1, "-12-31")
              
        for (day in 1:366){
                temp_table_yr_day <- filter(temp_table, V2 == day)
                at_df_bart_1$Site_ID[c] = "BART"
                at_df_bart_1$AT[c] = mean(temp_table_yr_day$V5, na.rm = TRUE)
                at_df_bart_1$ST_og[c] = mean(temp_table_yr_day$V6, na.rm = TRUE)
                at_df_bart_1$ST_cutoff[c] = mean(temp_table_yr_day$V7, na.rm = TRUE)
                at_df_bart_1$ST_20[c] = mean(temp_table_yr_day$V8, na.rm = TRUE)
                at_df_bart_1$ST_25[c] = mean(temp_table_yr_day$V9, na.rm = TRUE)
                at_df_bart_1$ST_30[c] = mean(temp_table_yr_day$V10, na.rm = TRUE)
                at_df_bart_1$ST_35[c] = mean(temp_table_yr_day$V11, na.rm = TRUE)
                at_df_bart_1$ST_40[c] = mean(temp_table_yr_day$V12, na.rm = TRUE)
                at_df_bart_1$ST_45[c] = mean(temp_table_yr_day$V13, na.rm = TRUE)
                at_df_bart_1$ST_50[c] = mean(temp_table_yr_day$V14, na.rm = TRUE)
                at_df_bart_1$ST_55[c] = mean(temp_table_yr_day$V15, na.rm = TRUE)
                at_df_bart_1$ST_60[c] = mean(temp_table_yr_day$V16, na.rm = TRUE)
                at_df_bart_1$ST_65[c] = mean(temp_table_yr_day$V17, na.rm = TRUE)
                at_df_bart_1$ST_70[c] = mean(temp_table_yr_day$V18, na.rm = TRUE)
                at_df_bart_1$ST_75[c] = mean(temp_table_yr_day$V19, na.rm = TRUE)
                at_df_bart_1$ST_80[c] = mean(temp_table_yr_day$V20, na.rm = TRUE)
                at_df_bart_1$ST_85[c] = mean(temp_table_yr_day$V21, na.rm = TRUE)
                at_df_bart_1$ST_90[c] = mean(temp_table_yr_day$V22, na.rm = TRUE)
                at_df_bart_1$ST_95[c] = mean(temp_table_yr_day$V23, na.rm = TRUE)
                at_df_bart_1$ST_100[c] = mean(temp_table_yr_day$V24, na.rm = TRUE)
              
                temp_table[V2 == year & V3 == day, mean(V6, na.rm = TRUE)]
                
                # need to convert date into "year-mo-day format"
                date <- as.Date(day, origin = org_date) 
                at_df_bart_1$Time[c] <- format(date, "%Y-%m-%d")
                c = c+1
                
         }
}

  # give soil T a year
soil_t$Year <- substr(soil_t$Time,1,4)
at_df_bart_1$Year <- substr(at_df_bart_1$Time,1,4)

# Make into date class
soil_t$Time <- as.Date(soil_t$Time)
at_df_bart_1$Time <- as.Date(at_df_bart_1$Time)

  # make a graph of a year of one of them
  temp_data <- filter(soil_t, Site_ID == matched$site_id[1])
  
  
  print(ggplot() +  
    geom_line(data = at_df_bart_1, aes(x = Time, y = ST_og ,group = 1, color = "Original Model ST"), linewidth = 1) +
    geom_line(data = temp_data, aes(x = Time, y = Neon_ST_column, group = 0.5, color = "NEON ST"), linewidth = 1) + 
      geom_line(data = at_df_bart_1, aes(x = Time, y = ST_55 ,group = 1, color = "Model ST with 55"), linewidth = 1) +
    geom_line(data = at_df_bart_1, aes(x = Time, y = ST_100 ,group = 1, color = "Model ST with 100 "), linewidth = 1) +
    geom_line(data = at_df_bart_1, aes(x = Time, y = ST_cutoff ,group = 1, color = "Model ST with Cutoff"), linewidth = 1) +
    labs(title = matched$site_id[1], x = "date", y = "Soil Temperature") + 
    scale_x_date(limits = as.Date(c("2017-01-01", "2018-12-31"))) +
    scale_color_manual(values = c("Original Model ST" = "orange", "Model ST with 100" = "pink", "Model ST with 55" = "red","Model ST with Cutoff" = "grey","NEON ST" = "lightgreen")))


```
Graph each new attempt
```{r}

# load in .clim file that i made

temp_table <- fread("ERA5.1.BART.sweep2.2012-01-01.2024-01-01.clim")
years = 2012:2024
length = years[length(years)]- years[1] + 3

at_df_bart <- data.table(
              Time = character(length*366),
              Site_ID = character(length*366),
              AT = numeric(length*366),
              ST_1 = numeric(length*366),
              ST_5 = numeric(length*366),
              ST_10 = numeric(length*366),
              ST_og = numeric(length*366),
              ST_20 = numeric(length*366),
              ST_cutoff = numeric(length*366)
            )
c = 1

for (year in years){
        temp_table_yr <- filter(temp_table, V1 == year)
              org_date = paste0(year - 1, "-12-31")
              
        for (day in 1:366){
                temp_table_yr_day <- filter(temp_table, V2 == day)
                at_df_bart$Site_ID[c] = "BART"
                at_df_bart$AT[c] = mean(temp_table_yr_day$V5, na.rm = TRUE)
                at_df_bart$ST_1[c] = mean(temp_table_yr_day$V6, na.rm = TRUE)
                at_df_bart$ST_cutoff[c] = mean(temp_table_yr_day$V7, na.rm = TRUE)
                at_df_bart$ST_5[c] = mean(temp_table_yr_day$V8, na.rm = TRUE)
                at_df_bart$ST_10[c] = mean(temp_table_yr_day$V9, na.rm = TRUE)
                at_df_bart$ST_og[c] = mean(temp_table_yr_day$V10, na.rm = TRUE)
                at_df_bart$ST_20[c] = mean(temp_table_yr_day$V11, na.rm = TRUE)
              
                temp_table[V2 == year & V3 == day, mean(V6, na.rm = TRUE)]
                
                # need to convert date into "year-mo-day format"
                date <- as.Date(day, origin = org_date) 
                at_df_bart$Time[c] <- format(date, "%Y-%m-%d")
                c = c+1
                
         }
}

  # give soil T a year
soil_t$Year <- substr(soil_t$Time,1,4)
at_df_bart$Year <- substr(at_df_bart$Time,1,4)

# Make into date class
soil_t$Time <- as.Date(soil_t$Time)
at_df_bart$Time <- as.Date(at_df_bart$Time)

  # make a graph of a year of one of them
  temp_data <- filter(soil_t, Site_ID == matched$site_id[1])
  
  
  print(ggplot() +  
    geom_line(data = at_df_bart, aes(x = Time, y = ST_og ,group = 1, color = "Original Model ST"), linewidth = 1) +
    geom_line(data = temp_data, aes(x = Time, y = Neon_ST_column, group = 0.5, color = "NEON ST"), linewidth = 1) + 
      geom_line(data = at_df_bart, aes(x = Time, y = ST_1 ,group = 1, color = "Model ST with 1"), linewidth = 1) +
  #  geom_line(data = at_df_bart, aes(x = Time, y = ST_10 ,group = 1, color = "Model ST with 10 "), linewidth = 1) +
    geom_line(data = at_df_bart, aes(x = Time, y = ST_cutoff ,group = 1, color = "Model ST with Cutoff"), linewidth = 1) +
    labs(title = matched$site_id[1], x = "date", y = "Soil Temperature") + 
    scale_x_date(limits = as.Date(c("2017-01-01", "2018-12-31"))) +
    scale_color_manual(values = c("Original Model ST" = "orange", "Model ST with 1" = "pink", "Model ST with 10" = "red","Model ST with Cutoff" = "grey","NEON ST" = "lightgreen")))


```



```{r}

# load in .clim file that i made

temp_table <- fread("ERA5.1.HARV.2012-01-01.2024-01-01.clim")
years = 2012:2024
length = years[length(years)]- years[1] + 3

at_df_harv <- data.table(
              Time = character(length*366),
              Site_ID = character(length*366),
              AT = numeric(length*366),
              ST_og = numeric(length*366),
              ST_20 = numeric(length*366),
              ST_25 = numeric(length*366),
              ST_30 = numeric(length*366),
              ST_35 = numeric(length*366),
              ST_40 = numeric(length*366),
              ST_45 = numeric(length*366),
              ST_50 = numeric(length*366),
              ST_55 = numeric(length*366),
              ST_60 = numeric(length*366),
              ST_65 = numeric(length*366),
              ST_70 = numeric(length*366),
              ST_75 = numeric(length*366),
              ST_80 = numeric(length*366),
              ST_85 = numeric(length*366),
              ST_90 = numeric(length*366),
              ST_95 = numeric(length*366),
              ST_100 = numeric(length*366),
              ST_cutoff = numeric(length*366)
            )
c = 1

for (year in years){
        temp_table_yr <- filter(temp_table, V1 == year)
              org_date = paste0(year - 1, "-12-31")
              
        for (day in 1:366){
                temp_table_yr_day <- filter(temp_table, V2 == day)
                at_df_bart$Site_ID[c] = "HARV"
                at_df_harv$AT[c] = mean(temp_table_yr_day$V5, na.rm = TRUE)
                at_df_harv$ST_og[c] = mean(temp_table_yr_day$V6, na.rm = TRUE)
                at_df_harv$ST_cutoff[c] = mean(temp_table_yr_day$V7, na.rm = TRUE)
                at_df_harv$ST_20[c] = mean(temp_table_yr_day$V8, na.rm = TRUE)
                at_df_harv$ST_25[c] = mean(temp_table_yr_day$V9, na.rm = TRUE)
                at_df_harv$ST_30[c] = mean(temp_table_yr_day$V10, na.rm = TRUE)
                at_df_harv$ST_35[c] = mean(temp_table_yr_day$V11, na.rm = TRUE)
                at_df_harv$ST_40[c] = mean(temp_table_yr_day$V12, na.rm = TRUE)
                at_df_harv$ST_45[c] = mean(temp_table_yr_day$V13, na.rm = TRUE)
                at_df_harv$ST_50[c] = mean(temp_table_yr_day$V14, na.rm = TRUE)
                at_df_harv$ST_55[c] = mean(temp_table_yr_day$V15, na.rm = TRUE)
                at_df_harv$ST_60[c] = mean(temp_table_yr_day$V16, na.rm = TRUE)
                at_df_harv$ST_65[c] = mean(temp_table_yr_day$V17, na.rm = TRUE)
                at_df_harv$ST_70[c] = mean(temp_table_yr_day$V18, na.rm = TRUE)
                at_df_harv$ST_75[c] = mean(temp_table_yr_day$V19, na.rm = TRUE)
                at_df_harv$ST_80[c] = mean(temp_table_yr_day$V20, na.rm = TRUE)
                at_df_harv$ST_85[c] = mean(temp_table_yr_day$V21, na.rm = TRUE)
                at_df_harv$ST_90[c] = mean(temp_table_yr_day$V22, na.rm = TRUE)
                at_df_harv$ST_95[c] = mean(temp_table_yr_day$V23, na.rm = TRUE)
                at_df_harv$ST_100[c] = mean(temp_table_yr_day$V24, na.rm = TRUE)
              
                temp_table[V2 == year & V3 == day, mean(V6, na.rm = TRUE)]
                
                # need to convert date into "year-mo-day format"
                date <- as.Date(day, origin = org_date) 
                at_df_harv$Time[c] <- format(date, "%Y-%m-%d")
                c = c+1
                
         }
}

  # give soil T a year
soil_t$Year <- substr(soil_t$Time,1,4)
at_df_harv$Year <- substr(at_df_harv$Time,1,4)

# Make into date class
soil_t$Time <- as.Date(soil_t$Time)
at_df_harv$Time <- as.Date(at_df_harv$Time)

  # make a graph of a year of one of them
  temp_data <- filter(soil_t, Site_ID == matched$site_id[2])
  
  
  print(ggplot() +  
    geom_line(data = at_df_harv, aes(x = Time, y = ST_og ,group = 1, color = "Original Model ST"), linewidth = 1) +
    geom_line(data = temp_data, aes(x = Time, y = Neon_ST_column, group = 0.5, color = "NEON ST"), linewidth = 1) + 
      geom_line(data = at_df_harv, aes(x = Time, y = ST_55 ,group = 1, color = "Model ST with 55"), linewidth = 1) +
    geom_line(data = at_df_harv, aes(x = Time, y = ST_100 ,group = 1, color = "Model ST with 100"), linewidth = 1) +
    geom_line(data = at_df_harv, aes(x = Time, y = ST_cutoff ,group = 1, color = "Model ST with Cutoff"), linewidth = 1) +
    labs(title = matched$site_id[2], x = "date", y = "Soil Temperature") + 
    scale_x_date(limits = as.Date(c("2017-01-01", "2018-12-31"))) +
    scale_color_manual(values = c("Original Model ST" = "orange", "Model ST with 100" = "pink", "Model ST with 55" = "red","Model ST with Cutoff" = "grey","NEON ST" = "lightgreen")))


```


MOAB
```{r}
# load in soil_t
load("soil_trm_data_12_8.RData")


# load in .clim file that i made

temp_table <- fread("ERA5.1.2012-01-01.MOAB.2024-01-01.clim")
years = 2012:2024
length = years[length(years)]- years[1] + 3

at_df_harv <- data.table(
              Time = character(length*366),
              Site_ID = character(length*366),
              AT = numeric(length*366),
              ST_bp = numeric(length*366),
              ST_exp = numeric(length*366),
              ST_bp_cutoff = numeric(length*366),
              ST_exp_cutoff = numeric(length*366)
            )
c = 1

#temp_table %>%
#  group_by("V2") %>%
#  summarize(average_value = mean(c = ("V3", "V4", "V5", "V6", "V7","V8", "V9"), na.rm = TRUE)

for (year in years){
        temp_table_yr <- filter(temp_table, V1 == year)
              org_date = paste0(year - 1, "-12-31")
              
        for (day in 1:366){
                temp_table_yr_day <- filter(temp_table, V2 == day)
                at_df_harv$Site_ID[c] = "HARV"
                at_df_harv$AT[c] = mean(temp_table_yr_day$V5, na.rm = TRUE)
                at_df_harv$ST_bp[c] = mean(temp_table_yr_day$V6, na.rm = TRUE)
                at_df_harv$ST_exp[c] = mean(temp_table_yr_day$V7, na.rm = TRUE)
                at_df_harv$ST_bp_cutoff[c] = mean(temp_table_yr_day$V8, na.rm = TRUE)
                at_df_harv$ST_exp_cutoff[c] = mean(temp_table_yr_day$V9, na.rm = TRUE)
          
                temp_table[V2 == year & V3 == day, mean(V6, na.rm = TRUE)]
                
                # need to convert date into "year-mo-day format"
                date <- as.Date(day, origin = org_date) 
                at_df_harv$Time[c] <- format(date, "%Y-%m-%d")
                c = c+1
                
         }
}

  # give soil T a year
soil_t$Year <- substr(soil_t$Time,1,4)
at_df_harv$Year <- substr(at_df_harv$Time,1,4)

# Make into date class
soil_t$Time <- as.Date(soil_t$Time)
at_df_harv$Time <- as.Date(at_df_harv$Time)

  # make a graph of a year of one of them
  temp_data <- filter(soil_t, Site_ID == matched$site_id[32])
  
  
  print(ggplot() +  
 #   geom_line(data = at_df_harv, aes(x = Time, y = ST_bp ,group = 1, color = "Original Model ST"), linewidth = 1) +
    geom_line(data = temp_data, aes(x = Time, y = Neon_ST_column, group = 0.5, color = "NEON ST"), linewidth = 1) + 
 #   geom_line(data = at_df_harv, aes(x = Time, y = ST_exp ,group = 1, color = "Model ST with Exp filter"), linewidth = 1) +
    geom_line(data = at_df_harv, aes(x = Time, y = ST_exp_cutoff ,group = 1, color = "Model ST with Exp filter & cutoff"), linewidth = 1) +
    geom_line(data = at_df_harv, aes(x = Time, y = ST_bp_cutoff ,group = 1, color = "Model ST with Cutoff"), linewidth = 1) +
    labs(title = matched$site_id[2], x = "date", y = "Soil Temperature") + 
    scale_x_date(limits = as.Date(c("2016-01-01", "2024-12-31"))) +
    scale_color_manual(values = c("Original Model ST" = "orange", "Model ST with Exp filter & cutoff" = "pink", "Model ST with Exp filter" = "red","Model ST with Cutoff" = "grey","NEON ST" = "lightgreen")))


```

HARV
```{r}

# load in soil_t
load("soil_trm_data_12_8.RData")


# load in .clim file that i made

temp_table <- fread("ERA5.1.2012-01-01.HARV.2024-01-01.clim")
years = 2012:2024
length = years[length(years)]- years[1] + 3

at_df_harv <- data.table(
              Time = character(length*366),
              Site_ID = character(length*366),
              AT = numeric(length*366),
              ST_bp = numeric(length*366),
              ST_exp = numeric(length*366),
              ST_bp_cutoff = numeric(length*366),
              ST_exp_cutoff = numeric(length*366)
            )
c = 1

#temp_table %>%
#  group_by("V2") %>%
#  summarize(average_value = mean(c = ("V3", "V4", "V5", "V6", "V7","V8", "V9"), na.rm = TRUE)

for (year in years){
        temp_table_yr <- filter(temp_table, V1 == year)
              org_date = paste0(year - 1, "-12-31")
              
        for (day in 1:366){
                temp_table_yr_day <- filter(temp_table, V2 == day)
                at_df_harv$Site_ID[c] = "HARV"
                at_df_harv$AT[c] = mean(temp_table_yr_day$V5, na.rm = TRUE)
                at_df_harv$ST_bp[c] = mean(temp_table_yr_day$V6, na.rm = TRUE)
                at_df_harv$ST_exp[c] = mean(temp_table_yr_day$V7, na.rm = TRUE)
                at_df_harv$ST_bp_cutoff[c] = mean(temp_table_yr_day$V8, na.rm = TRUE)
                at_df_harv$ST_exp_cutoff[c] = mean(temp_table_yr_day$V9, na.rm = TRUE)
          
                temp_table[V2 == year & V3 == day, mean(V6, na.rm = TRUE)]
                
                # need to convert date into "year-mo-day format"
                date <- as.Date(day, origin = org_date) 
                at_df_harv$Time[c] <- format(date, "%Y-%m-%d")
                c = c+1
                
         }
}

  # give soil T a year
soil_t$Year <- substr(soil_t$Time,1,4)
at_df_harv$Year <- substr(at_df_harv$Time,1,4)

# Make into date class
soil_t$Time <- as.Date(soil_t$Time)
at_df_harv$Time <- as.Date(at_df_harv$Time)

  # make a graph of a year of one of them
  temp_data <- filter(soil_t, Site_ID == matched$site_id[2])
  
  
  print(ggplot() +  
 #   geom_line(data = at_df_harv, aes(x = Time, y = ST_bp ,group = 1, color = "Original Model ST"), linewidth = 1) +
    geom_line(data = temp_data, aes(x = Time, y = Neon_ST_column, group = 0.5, color = "NEON ST"), linewidth = 1) + 
 #   geom_line(data = at_df_harv, aes(x = Time, y = ST_exp ,group = 1, color = "Model ST with Exp filter"), linewidth = 1) +
    geom_line(data = at_df_harv, aes(x = Time, y = ST_exp_cutoff ,group = 1, color = "Model ST with Exp filter & cutoff"), linewidth = 1) +
    geom_line(data = at_df_harv, aes(x = Time, y = ST_bp_cutoff ,group = 1, color = "Model ST with Cutoff"), linewidth = 1) +
    labs(title = matched$site_id[2], x = "date", y = "Soil Temperature") + 
    scale_x_date(limits = as.Date(c("2016-01-01", "2024-12-31"))) +
    scale_color_manual(values = c("Original Model ST" = "orange", "Model ST with Exp filter & cutoff" = "pink", "Model ST with Exp filter" = "red","Model ST with Cutoff" = "grey","NEON ST" = "lightgreen")))

```

KONZ
```{r}
# load in soil_t
load("soil_trm_data_12_8.RData")


# load in .clim file that i made

temp_table <- fread("ERA5.1.2012-01-01.KONZ.2024-01-01.clim")
years = 2012:2024
length = years[length(years)]- years[1] + 3

at_df_harv <- data.table(
              Time = character(length*366),
              Site_ID = character(length*366),
              AT = numeric(length*366),
              ST_bp = numeric(length*366),
              ST_exp = numeric(length*366),
              ST_bp_cutoff = numeric(length*366),
              ST_exp_cutoff = numeric(length*366)
            )
c = 1

#temp_table %>%
#  group_by("V2") %>%
#  summarize(average_value = mean(c = ("V3", "V4", "V5", "V6", "V7","V8", "V9"), na.rm = TRUE)

for (year in years){
        temp_table_yr <- filter(temp_table, V1 == year)
              org_date = paste0(year - 1, "-12-31")
              
        for (day in 1:366){
                temp_table_yr_day <- filter(temp_table, V2 == day)
                at_df_harv$Site_ID[c] = "HARV"
                at_df_harv$AT[c] = mean(temp_table_yr_day$V5, na.rm = TRUE)
                at_df_harv$ST_bp[c] = mean(temp_table_yr_day$V6, na.rm = TRUE)
                at_df_harv$ST_exp[c] = mean(temp_table_yr_day$V7, na.rm = TRUE)
                at_df_harv$ST_bp_cutoff[c] = mean(temp_table_yr_day$V8, na.rm = TRUE)
                at_df_harv$ST_exp_cutoff[c] = mean(temp_table_yr_day$V9, na.rm = TRUE)
          
                temp_table[V2 == year & V3 == day, mean(V6, na.rm = TRUE)]
                
                # need to convert date into "year-mo-day format"
                date <- as.Date(day, origin = org_date) 
                at_df_harv$Time[c] <- format(date, "%Y-%m-%d")
                c = c+1
                
         }
}

  # give soil T a year
soil_t$Year <- substr(soil_t$Time,1,4)
at_df_harv$Year <- substr(at_df_harv$Time,1,4)

# Make into date class
soil_t$Time <- as.Date(soil_t$Time)
at_df_harv$Time <- as.Date(at_df_harv$Time)

  # make a graph of a year of one of them
  temp_data <- filter(soil_t, Site_ID == matched$site_id[15])
  
  
  print(ggplot() +  
 #   geom_line(data = at_df_harv, aes(x = Time, y = ST_bp ,group = 1, color = "Original Model ST"), linewidth = 1) +
    geom_line(data = temp_data, aes(x = Time, y = Neon_ST_column, group = 0.5, color = "NEON ST"), linewidth = 1) + 
 #   geom_line(data = at_df_harv, aes(x = Time, y = ST_exp ,group = 1, color = "Model ST with Exp filter"), linewidth = 1) +
    geom_line(data = at_df_harv, aes(x = Time, y = ST_exp_cutoff ,group = 1, color = "Model ST with Exp filter & cutoff"), linewidth = 1) +
    geom_line(data = at_df_harv, aes(x = Time, y = ST_bp_cutoff ,group = 1, color = "Model ST with Cutoff"), linewidth = 1) +
    labs(title = matched$site_id[2], x = "date", y = "Soil Temperature") + 
    scale_x_date(limits = as.Date(c("2016-01-01", "2024-12-31"))) +
    scale_color_manual(values = c("Original Model ST" = "orange", "Model ST with Exp filter & cutoff" = "pink", "Model ST with Exp filter" = "red","Model ST with Cutoff" = "grey","NEON ST" = "lightgreen")))

```

