---
title: "soilmoist_validation"
author: "chaney"
date: "2025-11-24"
output: html_document
---
```{r setup, include=FALSE}
# Load libraries

library(DBI)
library(dplyr)
library(geosphere)
library(sp)
library(ggplot2)
library(sf)
library(maps)
library(terra)
library(knitr)
library(tidyr)
library(neonUtilities)
library(swfscMisc)
library(maps)
library(tidyverse)
library(neonSoilFlux)
library(parallel)
library(ncdf4)
library(data.table)
library(purrr)
library(data.table)
library(parallel)
library(lubridate)
library(terra)
library(httr)
library(sp)
library(neonSoilFlux)
library(stringr)
library(scales)
library(MASS)
library(neonstore)

```

---- First, get all of the data -----

Find which model sites correspond to NEON sites
```{r}
# get the coordinates of 6400 sites in Dongchen's output
Neonplots <- read.csv("/usr4/ugrad/chaney/R/3_25/Neon_sites_terrestrial.csv") 

# Add a column for site numbers in Site_Info 
Site_Info <- readRDS("/usr4/ugrad/chaney/R/3_25/site.locs.rds")
Site_Info$Site_Number <- seq(1, nrow(Site_Info)) 

# Extract latitude and longitude from Neonplots and Site_Info tables
neon_coords <- Neonplots[ ,2:3] %>%
  select(Latitude = field_latitude, Longitude = field_longitude)
Site_Info_coords <- Site_Info %>%
  select(Site_Number, lat, lon)

# Put Site_Info latitude/longitude in a matrix
Site_Coordinates <- as.matrix(Site_Info[, c("lat", "lon")])

# Same for Neonplots
Neonplots_Coordinates <- as.matrix(Neonplots[, 2:3])

# Make list to store matched list
matched <- data.frame(id = numeric(nrow(Neonplots)), # id in forecast
                      site_id = numeric(nrow(Neonplots))) # NEON id

# Calculate each plot's closest site number
for (i in 1:nrow(Neonplots)) {
  # Extract the coordinates for the current row in Neonplots
  plot_coords <- Neonplots_Coordinates[i, ]
  
  # Compute distances from this plot to all site coordinates (longlat means its in km)
  distances <- spDistsN1(Site_Coordinates, plot_coords, longlat = TRUE)
  
  # Find the index of the minimum distance for closest site
  closest_site_index <- which.min(distances)
  
  # Assign the corresponding Site_Number from Site_Info to the current row in Neonplots
  matched$id[i] <- Site_Info$Site_Number[closest_site_index]
  matched$site_id[i] <- Neonplots$field_site_id[i]
}


```


Get soil moisture data from model -- Percent of Soil Saturation
```{r}
source("neon_model_processing.R")

# Set outdir
outdir <- "/projectnb/dietzelab/dongchen/anchorSites/NA_runs/SDA_8k_site/NA_SDA_no_Debias";

final_daily_sm_data <- neon_model_processing(matched, 2012, 2025, outdir, "SoilMoistFrac", var2 = NULL)

colnames(final_daily_sm_data)[colnames(final_daily_sm_data) == "Output1"] <- "Model_SM"
```


Get NEON Soil moisture from John's data in SCC -- Volumetric Soil Water Content
```{r}

# Function to extract John's data
soil_moist <- function(site_number){
  
  years = 2012:2025
  site = matched$site_id[site_number]
  
  # create list to store data
  all_days <- list() 
  count <- 1 


  for (year in years){
    for (i in 01:12){
      if (i < 10){
        file_name <- paste0("/projectnb/dietzelab/jzobitz/02-NEON-sites/env-data/env-meas-", site, "-",year,"-0",i,".Rda")
      }else{
        file_name <- paste0("/projectnb/dietzelab/jzobitz/02-NEON-sites/env-data/env-meas-", site, "-",year,"-",i,".Rda")
      }
      if (file.exists(file_name)== TRUE){
        load(file_name)
        
        # Get VSWC content data
        site_data_moist <- site_data[[2]][[2]] #V water/ V total soil
        
        # Sort by date
        site_data_moist <- site_data_moist[order(site_data_moist$startDateTime),]
        
        # Combine positions into one column 
        site_data_moist$Position <- paste(site_data_moist$horizontalPosition,site_data_moist$verticalPosition)
        
        # Find the number of data points per day (48 data points at multiple data sites and depths for each site)
        dppd = 48 * length(unique(site_data_moist$Position))
        
        # Find the number of days 
        n_days <- nrow(site_data_moist) / (dppd)
        
        # Add day index 
        site_data_moist$day_idx <- ceiling(seq_len(nrow(site_data_moist)) / dppd)
      
       # Count rows per day
        day_sizes <- table(site_data_moist$day_idx)
        
        # Keep only complete days
        complete_days <- as.numeric(names(day_sizes[day_sizes == dppd]))
        
        filtered <- site_data_moist %>% 
          filter(day_idx %in% complete_days)
          
        # Turn VSWC into column where column = day
        vswc_mat <- matrix(filtered$VSWCMean, nrow = dppd, ncol = length(complete_days))
      
        # Find daily means VSWC
        daily_means <- colMeans(vswc_mat, na.rm = TRUE)
        
        # Find times
       day_times <- filtered$startDateTime[match(complete_days, filtered$day_idx)]
        
        # Find largest verticalPosition
        deepest_vert <- max(unique(site_data_moist$verticalPosition))
        
        # Filter for just that vert position
        deepest_rows <- filter(site_data_moist, verticalPosition == deepest_vert)
        
        # Find average depth of deepest_rows (per month)
        depth <- mean(deepest_rows$zOffset, na.rm = TRUE)
        
        # Add to table
        summary_table <- data.frame(
          Site_ID = site,
          Time = day_times,
          Soil_moist_neon = daily_means,
          Ave_depth = depth
        )
        
        # Add table to list
      all_days[[count]] <- summary_table
      count <- count + 1
        
      }
    }
  }
   # turn list into table
  summary <- rbindlist(all_days)
  
return(list(stm = summary))
}


# Apply parallel calculation for each site
n = 1:nrow(matched)
results_m <- mclapply(n, soil_moist, mc.cores = num_cores)

# Combine results from all ensembles
soil_m <- rbindlist(lapply(results_m, `[[`, "stm"),fill=TRUE)

# Delete all rows with just zeros
soil_m <- filter(soil_m, Soil_moist_neon != 0)

```


NEON data is VSWC, so needs to be converted to Percent Soil Saturation. This can be done by dividing VSWC by porosity.

Porosity can be calculated using NEON megapit data and function from Pecan Github
```{r}
start_date = NA
end_date = NA 
store_dir <- "/Users/chane/OneDrive/Documents/ef_lab/R_files" #

# Load function from Pecan
source("soil_utils.R")

# Make df to keep everything
neon_soil_texture <- data.frame(Site_ID = character(length(matched$site_id)),
                              percentSand = numeric(length(matched$site_id)),
                              percentSilt = numeric(length(matched$site_id)),
                              percentClay = numeric(length(matched$site_id)),
                              porosity = numeric(length(matched$site_id)))

# Loop through all sites and get NEON data
for(i in 1:length(matched$site_id)){
  
  neon_soil_texture$Site_ID[i] <- matched$site_id[i]
  
  # get neon Data
neonstore::neon_download("DP1.00096.001", dir = store_dir, table = NA, site = matched$site_id[i], start_date = NA, end_date = NA, type = "basic",api = "https://data.neonscience.org/api/v0")

# Contains horizon depths
  perarchivesample <- neonstore::neon_read(table = "perarchivesample", product = "DP1.00096.001", site = matched$site_id[i], start_date = NA, end_date = NA, dir = store_dir)
  
  # Contains soil horizon percent sand/silt/clay 
  perbiogeosample <- neonstore::neon_read(table = "perbiogeosample", product = "DP1.00096.001", site = matched$site_id[i], start_date = NA, end_date = NA, dir = store_dir)
  
  # Merge archive and biogeo samples 
  persample <- merge(perarchivesample, perbiogeosample, by = c("domainID","siteID","horizonName", "pitID","horizonID","pitNamedLocation"))
  
    # Sort by depth
  persample<-persample[order(persample$archiveTopDepth),]
  
    # Delete rows where sandTotal, siltTotal, and clayTotal are na
  for (row in 1:nrow(persample)){
    if (is.na(persample$sandTotal[row]) == TRUE & is.na(persample$siltTotal[row]) == TRUE & is.na(persample$clayTotal[row]) == TRUE){
      persample <- persample[-row, ]
    }
  }
  
  # -- Find row with depth closest to average depth to depth from VSWC
  
  # Find rows of correct site
  row <- which(soil_m$Site_ID == matched$site_id[i]) 
  
  # If this site exists, find porosity
  if ((row[1] > 0) == TRUE){
 
     # average across rows and multiply by 100 since in m (archiveBottomDepth is in cm)
    neon_depth <- (mean(soil_m$Ave_depth[row], na.rm = TRUE))*100 # cm

   # Calculate absolute differences between target and each neon depth
    abs_diff <- abs(persample$archiveBottomDepth + neon_depth) # NEON depth negative
    
    # Find the index of the minimum absolute difference
    closest_row_index <- which.min(abs_diff)
    
  # -- Find weighted average of rows above and including the closest row 
    horizon_length <- persample$archiveBottomDepth - persample$archiveTopDepth # cm
    
    # Find the average sand, silt, and clay, weighted by how deep the layer is
    
    # set to zero
    sand_ave <- 0
    silt_ave <- 0
    clay_ave <- 0
    
    
    for (horizon in 1:closest_row_index){
      sand_ave <- sand_ave + (((horizon_length[horizon])/(sum(horizon_length[1:closest_row_index])))*persample$sandTotal[horizon])
      silt_ave <- silt_ave + (((horizon_length[horizon])/(sum(horizon_length[1:closest_row_index])))*persample$siltTotal[horizon])
      clay_ave <- clay_ave + (((horizon_length[horizon])/(sum(horizon_length[1:closest_row_index])))*persample$clayTotal[horizon])
    }
  
  neon_soil_texture$percentSand[i] <- sand_ave
  neon_soil_texture$percentSilt[i] <- silt_ave
  neon_soil_texture$percentClay[i] <- clay_ave
  
  # Use soil_params function to get porostiy from silt,sand,clay percentages
    if (is.na(mean(neon_soil_texture$percentSand[i])) == TRUE){
      neon_soil_texture$porosity[i] = NA
    }else{
        mysoil <- soil_params(soil_type=NULL, sand=neon_soil_texture$percentSand[i], silt=neon_soil_texture$percentSilt[i], clay=neon_soil_texture$percentClay[i], bulk=NULL)
        neon_soil_texture$porosity[i] <- mysoil$volume_fraction_of_water_in_soil_at_saturation
    }
  }
}
```

Divide VSWC by porosity to get soil saturation
```{r}

# Add a row percent soil saturation
soil_m$Neon_SM <- numeric(length(soil_m$Time))

# Loop through each site and calculate Soil saturation
for (i in 1:length(soil_m$Site_ID)){
      row <- which(neon_soil_texture$Site_ID == soil_m$Site_ID[i])
      
      if (is.na(neon_soil_texture$porosity[row]) == FALSE){
        soil_m$Neon_SM[i]= soil_m$Soil_moist_neon[i]/neon_soil_texture$porosity[row]
      }else{
        soil_m$Neon_SM[i] = NA
      }
}
```


Combine everything into one table 
```{r}

# Delete var2
final_daily_sm_data$Output2 <- NULL

# Change Time type 
soil_m$Time <- as.Date(soil_m$Time)
final_daily_sm_data$Time <- as.Date(final_daily_sm_data$Time)

# Combine two soil temperature tables
soil_m_table <- merge(final_daily_sm_data, soil_m, by = c("Time","Site_ID"))

load("/Users/chane/Documents/soil_table.Rdata") # load in table with soil respiration data (from "soilrespiration_validation.Rmd")

# Merge with soil respiration 
soil_table <- merge(soil_table, soil_m_table, by = c("Time","Site_ID"))

# Change some names
names(soil_m_table)[names(soil_m_table) == "soil_sat_neon"] <- "Neon_SM"
names(soil_m_table)[names(soil_m_table) == "Soil_moist_model"] <- "Model_SM"

names(soil_table)[names(soil_table) == "soil_sat_neon"] <- "Neon_SM"
names(soil_table)[names(soil_table) == "Soil_moist_model"] <- "Model_SM"


```


---- Next, Look at NEON relationship of Soil R as a function of Soil M 

Graph with data
```{r}
# Loess Regressions 
df_m_ma <- data.frame(x = soil_table$Neon_SM,
                   y = soil_table$Marshall_SR)
df_m_ma <- df_m_ma %>%
           filter(if_all(where(is.numeric), is.finite))

df_m_mq <- data.frame(x = soil_table$Neon_SM,
                   y = soil_table$Mill_Quirk_SR)
df_m_mq <- df_m_mq %>%
           filter(if_all(where(is.numeric), is.finite))

loess90_m_ma <- loess(y ~ x, data=df_m_ma, span=.9, na.action = na.omit)
df_m_ma$y_fit <- predict(loess90_m_ma)

loess90_m_mq <- loess(y ~ x, data=df_m_mq, span=.9)
df_m_mq$y_fit <- predict(loess90_m_mq)

# Graph
ggplot() +
  geom_point(data = df_m_ma, aes(x = x, y = y, color = "Marshall data"), size = 1) +
  geom_point(data = df_m_mq, aes(x = x, y = y, color = "Millington Quirk data"), size = 1) +
  geom_line(data = df_m_ma, aes(x = x, y = y_fit, color = "Marshall Loess Regression"), size = 1) +
    geom_line(data = df_m_mq, aes(x = x, y = y_fit, color = "Millington Quirk Loess Regression"), size = 1) +
  scale_color_manual(values = c("Marshall Data" = "darkgrey", "Millington Quirk Data" = "lightgrey", "Marshall Loess Regression" = "darkgreen", "Millington Quirk Loess Regression" = "lightgreen")) +
    labs(title = "Regression Model for Soil R as a function of Moisture",
       x = "Percent Soil Saturation",
       y = "Soil Respiration",
       color = "Dataset")


```

Line fits for each NEON site Moist
```{r}

# Find unique site names
sites <- unique(soil_table$Site_ID)

# Fit a line for each site
for (i in 1:length(sites)){
  temp_data <- filter(soil_table, Site_ID == sites[i])
  
  temp_data <- temp_data[order(temp_data$Neon_SM)]
  
  df <- data.frame(x = temp_data$Neon_SM,
                   y = temp_data$Marshall_SR)

  # Scatter Plot with Fitted line
  print(ggplot(df, aes(x = x, y = y)) + geom_point(size = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = sites[i], x = "Soil Moisture", y = "Soil Respiration"))

}
```

Fitting a parabola to Soil R as a function of SM
```{r}
for (i in 1:length(sites)){
  
  # Sort Data
  temp_data <- filter(soil_table, Site_ID == sites[i])
  temp_data <- temp_data[order(temp_data$Neon_SM)]
  
  # Get data frame
  df_temp_ma <- data.frame(x = temp_data$Neon_SM,
                   y = temp_data$Marshall_SR)
  
   df_temp_ma <- df_temp_ma %>%
           filter(if_all(where(is.numeric), is.finite))
   
  df_temp_mq <- data.frame(x = temp_data$Neon_SM,
                   y = temp_data$Mill_Quirk_SR)
   df_temp_mq <- df_temp_mq %>%
           filter(if_all(where(is.numeric), is.finite))
  
  
  if (is.na(df_temp_mq$x[1]) == FALSE){

    # Loess Regression
    loess90_temp_ma <- loess(y ~ x, data=df_temp_ma, span=.9)
    df_temp_ma$y_fit_loess <- predict(loess90_temp_ma)
  
    loess90_temp_mq <- loess(y ~ x, data=df_temp_mq, span=.9)
    df_temp_mq$y_fit_loess <- predict(loess90_temp_mq)
    
  
   # Parabola fit 
    parabola_fit_ma_lm <- lm(y ~ x + I(x^2), data = df_temp_ma, na.action = na.exclude)
 df_temp_ma$y_fit_parabola_lm <- predict(parabola_fit_ma_lm)

  parabola_fit_mq_lm <- lm(y ~ x + I(x^2), data = df_temp_mq, na.action = na.exclude)
  df_temp_mq$y_fit_parabola_lm <- predict(parabola_fit_mq_lm)
  
    # Parabola fit with outliers weighted less (looks same)
    parabola_fit_ma_rlm <- rlm(y ~ x + I(x^2), data = df_temp_ma, na.action = na.exclude)
 df_temp_ma$y_fit_parabola_rlm <- predict(parabola_fit_ma_rlm)

  parabola_fit_mq_rlm <- rlm(y ~ x + I(x^2), data = df_temp_mq, na.action = na.exclude)
  df_temp_mq$y_fit_parabola_rlm <- predict(parabola_fit_mq_rlm)
  
  
  
     # Graph Loess and Parabola for each site
    print(ggplot() +
      geom_point(data =  df_temp_ma, aes(x = x, y = y, color = "Marshall Data"), size = 1) +
    geom_point(data =  df_temp_mq, aes(x = x, y = y, color = "Mill Qurik Data"), size = 1) +
      geom_line(data =  df_temp_ma, aes(x = x, y = y_fit_loess, color = "Marshall Loess Fit"), size = 1) +
     geom_line(data =  df_temp_mq, aes(x = x, y = y_fit_loess, color = "Mill Quirk Loess Fit"), size = 1) +
 #    geom_line(data =  df_temp_ma, aes(x = x, y = y_fit_parabola_lm, color = "Marshall Parabolic Fit LM"), size = 1) +
 #    geom_line(data =  df_temp_mq, aes(x = x, y = y_fit_parabola_lm, color = "Mill Quirk Parabolic Fit LM"), size = 1) +
           geom_line(data =  df_temp_ma, aes(x = x, y = y_fit_parabola_rlm, color = "Marshall Parabolic Fit RLM"), size = 1) +
     geom_line(data =  df_temp_mq, aes(x = x, y = y_fit_parabola_rlm, color = "Mill Quirk Parabolic Fit RLM"), size = 1) +
       scale_color_manual(values = c("Marshall Data" = "darkgrey", "Mill Quirk Data" = "lightgrey", "Marshall Loess Fit" = "darkgreen", "Mill Quirk Loess Fit" = "lightgreen", "Marshall Parabolic Fit LM" = "royalblue", "Mill Quirk Parabolic Fit LM" = "lightblue", "Marshall Parabolic Fit RLM" = "purple", "Mill Quirk Parabolic Fit RLM" = "lavender")) +
      labs(title = sites[i],
           x = "Soil Moisture",
           y = "Soil Respiration",
           color = "Dataset"))
  }

}


```

Time series
```{r}
for (i in 1:nrow(matched)){

  temp_data <- filter(soil_table, Site_ID == matched$site_id[i])
  
  # Look at just one year for now
  temp_data_2022 <- filter(temp_data, Year == "2022")
  
  print(ggplot(temp_data_2022) +  
    geom_line(aes(x = Time, y = Model_SM,group = 1), linewidth = 1, color = "orange") +
    geom_line(aes(x = Time, y = Neon_SM, group = 1), linewidth = 1, color = "darkgreen") + 
    labs(title = matched$site_id[i], x = "date", y = "Percent Soil Saturation"))
}
```

Get precipitation data from NEON, so we can compare to SM data
```{r}

# Get NEON precipitation (tipping bucket) for all sites
neon_download(product = "DP1.00045.001", site = NA, type = "basic")

# Extract data for every 30 min
precip_data <- neon_read(product = "DP1.00045.001", table = "TIPPRE_30min")

# Add a day and year column
precip_data$Year <- substr(precip_data$startDateTime,1,4)
precip_data$Day <- substr(precip_data$startDateTime,1,10)

# Change sitename to match
colnames(precip_data)[colnames(precip_data) == "siteID"] <- "Site_ID"


# Aggregate to everyday
  precip_data_day <- precip_data %>%
  group_by(Day, Site_ID) %>%
  summarize(across(c(precipBulk,Year), mean, na.rm = TRUE), .groups = 'drop')
colnames(precip_data_day)[colnames(precip_data_day) == "Day"] <- "Time"

precip_data_day$Year <- substr(precip_data_day$Day,1,4)

```

Graph Precipitation of each site with SM data (for one year)
```{r}
for (i in 1:nrow(matched)){
  
  temp_data_sm <- filter(soil_table, Site_ID == matched$site_id[i])
  
  # Look at just one year for now
  temp_data_sm_2022 <- filter(temp_data_sm, Year == "2022")
  
  # Sort prec data
  temp_data_prec <- filter(precip_data_day,  Site_ID == matched$site_id[i])
  temp_data_prec_2022 <- filter(temp_data_prec,  Year == "2019")
  
  # Normalize precip data
  temp_data_prec_2022$precipBulk_norm <- temp_data_prec_2022$precipBulk/ max(temp_data_prec_2022$precipBulk)
  
 print(ggplot() +  
    geom_line(data = temp_data_prec_2022, aes(x = as.Date(Day), y = precipBulk_norm, group = 1, color = "Normalized Precipitation"), linewidth = 1)+
    geom_line(data = temp_data_sm_2022, aes(x = as.Date(Time), y = Model_SM,group = 1, color = "Model"), linewidth = 1) +
    geom_line(data = temp_data_sm_2022, aes(x = as.Date(Time), y = Neon_SM, group = 1, color = "NEON"), linewidth = 1) + 
    labs(title = matched$site_id[i], x = "date", y = "Percent Soil Saturation") + 
   scale_color_manual(values = c("Normalized Precipitation" = "black", "Model" = "orange", "NEON" = "lightgreen")))
 
}
```


Look at one month of precipitation data 
```{r}

# Add month column
precip_data_day$Month <- substr(precip_data_day$Day,1,7)
soil_table$Month <- substr(soil_table$Time,1,7)

for (i in 1:nrow(matched)){

  temp_data_sm <- filter(soil_table, Site_ID == matched$site_id[i])
  
  # Look at just one month for now
  temp_data_sm_2018_4 <- filter(temp_data_sm, Month == "2018-04")
  
  # sort prec data
  temp_data_prec <- filter(precip_data_day,  Site_ID == matched$site_id[i])
  temp_data_prec_2018_04 <- filter(temp_data_prec,  Month == "2018-04")
  
  # Normalize precip data
  temp_data_prec_2018_04$precipBulk_norm <- temp_data_prec_2018_04$precipBulk/ max(temp_data_prec_2018_04$precipBulk)
  
 print(ggplot() +  
    geom_line(data = temp_data_prec_2018_04, aes(x = as.Date(Day), y = precipBulk_norm, group = 1, color = "Normalized Precipitation"), linewidth = 1)+
    geom_line(data = temp_data_sm_2018_04, aes(x = as.Date(Time), y = Model_SM,group = 1, , color = "Model"), linewidth = 1) +
    geom_line(data = temp_data_sm_2018_04, aes(x = as.Date(Time), y = Neon_SM, group = 1, color = "NEON"), linewidth = 1) + 
      labs(title = matched$site_id[i], x = "date", y = "Percent Soil Saturation") + 
    scale_color_manual(values = c("Normalized Precipitation" = "black", "Model" = "orange", "NEON" = "lightgreen")))
 
}
```

Graphing each year for all sites
```{r}
for (y in 2015:2025){
  
  # Sort by year
  temp_data_sm <- filter(soil_table, Site_ID == matched$site_id[i])
  
  # Sort by Site
  temp_data_sm_2022 <- filter(temp_data_sm, Year == y)
  
  # Sort by Prec data
  temp_data_prec <- filter(precip_data_day,  Site_ID == matched$site_id[i])
  temp_data_prec_2022 <- filter(temp_data_prec,  Year == y)
  
  # Normalize precip data
  temp_data_prec_2022$precipBulk_norm <- temp_data_prec_2022$precipBulk/ max(temp_data_prec_2022$precipBulk)
  
 print(ggplot() +  
    geom_line(data = temp_data_prec_2022, aes(x = as.Date(Day), y = precipBulk_norm, group = 1, color = "Normalized Precipitation"), linewidth = 1)+
    geom_line(data = temp_data_sm_2022, aes(x = as.Date(Time), y = Model_SM,group = 1, , color = "Model"), linewidth = 1) +
    geom_line(data = temp_data_sm_2022, aes(x = as.Date(Time), y = Neon_SM, group = 1, color = "NEON"), linewidth = 1) + 
      labs(title = matched$site_id[i], x = "date", y = "Percent Soil Saturation") + 
    scale_color_manual(values = c("Normalized Precipitation" = "black", "Model" = "orange", "NEON" = "lightgreen")))
 
}
```


---- Now, Validating Model ---- 

-- Basic Error Validations 

Predicted-Observed Plots for each site
```{r}

soil_moist_line_fits<- data.frame(Site_ID = character(length(matched$site_id)),
                                slope = numeric(length(matched$site_id)),
                                intercept = numeric(length(matched$site_id)))


# Make predicted- observed plots
for (i in 1:length(sites)){
  temp_data <- filter(soil_table, Site_ID == sites[i])
  
  print(ggplot(temp_data, aes(x = Model_SM, y = Neon_SM)) + geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
        geom_abline(slope = 1, intercept = 0, color = "red", size = 2)+
  labs(title = sites[i], x = "Predicted SM", y = "Observed SM"))
  
   # Fit line
  line_fit <- lm(Neon_SM~Model_SM, data = temp_data, na.action = na.exclude)
  coefs<- coef(line_fit)
  
  # Add to table
  soil_moist_line_fits$Site_ID[i] = sites[i];
  soil_moist_line_fits$slope[i] = unname(coefs["(Intercept)"]);
  soil_moist_line_fits$intercept[i] = unname(coefs["Soil_temp_model"]);

}
```

Finding RMS error and correlation coefficients
```{r}

# Calculate error
soil_table$SMoist_error <- soil_table$Model_SM - soil_table$Neon_SM

# Keep a permanent daily_sr
soil_table_perm <- soil_table

# Make table for summary statistics
summary_stats_M <- data.frame(Site_ID = character(length(sites)),
                           SMoist_corr_coef = numeric(length(sites)),
                           SMoist_rms_error = numeric(length(sites)))


# Do summary stats
sites <- unique(soil_table$Site_ID)

for (i in 1:length(sites)){
  temp_data <- filter(soil_table, Site_ID == sites[i])
  
  # Remove na and inf values
  temp_data <- temp_data %>%
           filter(if_all(where(is.numeric), is.finite))
  
  # Add site name
  summary_stats_M$Site_ID[i] <- sites[i]
  
  # Find correlation coeffs
  summary_stats_M$SMoist_corr_coef[i] <- cor(temp_data$Neon_SM, temp_data$Model_SM)
  
  # Find RMS errors
  summary_stats_M$SMoist_rms_error[i] <- sqrt(mean((temp_data$Neon_SM - temp_data$Model_SM)^2))
}

```


Graph summary stats
```{r}
# Add site coordinates
summary_stats_M$Site_Lat <- numeric(length(sites))
summary_stats_M$Site_Long <- numeric(length(sites))

for (i in 1:length(summary_stats_M$Site_ID)){
  site_num = which(Neonplots$field_site_id == summary_stats_M$Site_ID[i])
  summary_stats_M$Site_Lat[i] <- Neonplots$field_latitude[site_num]
  summary_stats_M$Site_Long[i] <- Neonplots$field_longitude[site_num]
}


USA = map_data("world") %>% filter(region == "USA")
us_map <- st_as_sf(USA, 
            coords = c("long", "lat"), 
            crs = 4326)  # CRS 4326 is WGS84 (latitude/longitude)

# Map of correlation coefficient

print(ggplot()+
  geom_polygon(data = USA, aes(x= long, y = lat, group = group), fill="forestgreen", alpha = 0.3) + geom_point(data = summary_stats_M, aes(x= Site_Long, y = Site_Lat, size = SMoist_corr_coef, color = SMoist_corr_coef)) + geom_point(alpha = 0.7) + xlim(-170, -50) + scale_size_continuous(range = c(1,10)) + theme_void() + labs(title="Soil Moisture Coefficient") + theme(plot.title = element_text(hjust = 0.5),))


```




-- Next, validating Soil R as a function of Soil M relationship 


Perform Loess regressions to fit curve of soil respiration as a function of soil moisture
```{r}

# LOESS Regressions for model
df_m_model <- data.frame(x = soil_table$Model_SM,
                   y = soil_table$Model_SR)
df_m_model <- df_m_model %>%
           filter(if_all(where(is.numeric), is.finite))

loess90_m_model <- loess(y ~ x, data=df_m_model, span=.9)
df_m_model$y_fit <- predict(loess90_m_model)

ggplot() +
  geom_line(data = df_m_ma, aes(x = x, y = y_fit, color = "NEON Data: Marshall Method"), size = 1) +
    geom_line(data = df_m_mq, aes(x = x, y = y_fit, color = "NEON Data: Mill Quirk Method"), size = 1) +
  geom_line(data = df_m_model, aes(x = x, y = y_fit, color = "Model Output"), size = 1)+
  scale_color_manual(values = c("Model Output" = "orange", "NEON Data: Mill Quirk Method" = "lightgreen", "NEON Data: Marshall Method" = "darkgreen")) +
    labs(title = "Regression Model for Soil R as a function of M",
       x = "Soil Moisture",
       y = "Soil Respiration",
       color = "Dataset")

```

LOESS Regression and Parabolic Fit for NEON & Model at each site 
```{r}

# Make a list to store loess and parabolic fits
SM_fits <- list()

# Table to keep track of vertices
vertices <- data.frame(Site_ID = character(length(sites)),
                                  vertex_x_ma = numeric(length(sites)),
                                  vertex_x_mq = numeric(length(sites)),
                                  vertex_x_model = numeric(length(sites)),
                                  vertex_y_ma = numeric(length(sites)),
                                  vertex_y_mq = numeric(length(sites)),
                                  vertex_y_model = numeric(length(sites)))

for (i in 1:length(sites)){
  temp_data <- filter(soil_table, Site_ID == sites[i])
  
  temp_data <- temp_data[order(temp_data$Neon_SM),]
  
  df_temp_ma <- data.frame(Time = temp_data$Time,
                     Site_ID = temp_data$Site_ID,
                     Neon_SM = temp_data$Neon_SM,
                   Marshall_SR = temp_data$Marshall_SR)
  df_temp_ma <- df_temp_ma %>%
           filter(if_all(where(is.numeric), is.finite))
  
  df_temp_mq <- data.frame(Time = temp_data$Time,
                     Site_ID = temp_data$Site_ID,
                     Neon_SM = temp_data$Neon_SM,
                   Mill_Quirk_SR = temp_data$Mill_Quirk_SR)
  df_temp_mq <- df_temp_mq %>%
           filter(if_all(where(is.numeric), is.finite))
  
  df_temp_model <- data.frame(Time = temp_data$Time,
                     Site_ID = temp_data$Site_ID,
                     Model_SM = temp_data$Model_SM,
                   Forecast_SR = temp_data$Model_SR)
  df_temp_model <- df_temp_model %>%
           filter(if_all(where(is.numeric), is.finite))
  
  if (is.na(df_temp_mq$Neon_SM[1]) == FALSE){

    # LOESS 
    # NEON
    loess90_temp_ma <- loess(Marshall_SR ~ Neon_SM, data=df_temp_ma, span=.9)
    df_temp_ma$y_fit_loess_ma <- predict(loess90_temp_ma)
  
    loess90_temp_mq <- loess(Mill_Quirk_SR ~ Neon_SM, data=df_temp_mq, span=.9)
    df_temp_mq$y_fit_loess_mq <- predict(loess90_temp_mq)
    
    # SDA Model
    loess90_temp_model <- loess(Forecast_SR ~ Model_SM, data=df_temp_model, span=.9)
    df_temp_model$y_fit_loess_model <- predict(loess90_temp_model)
    
      # Parabola fit 
    parabola_fit_ma_rlm <- rlm(Marshall_SR ~ Neon_SM + I(Neon_SM^2), data = df_temp_ma, na.action = na.exclude)
    df_temp_ma$y_fit_parabola_rlm_ma <- predict(parabola_fit_ma_rlm)

     parabola_fit_mq_rlm <- rlm(Mill_Quirk_SR ~ Neon_SM + I(Neon_SM^2), data = df_temp_mq, na.action = na.exclude)
     df_temp_mq$y_fit_parabola_rlm_mq <- predict(parabola_fit_mq_rlm)
     
     parabola_fit_model_rlm <- rlm(Forecast_SR ~ Model_SM + I(Model_SM^2), data = df_temp_model, na.action = na.exclude)
     df_temp_model$y_fit_parabola_rlm_model <- predict(parabola_fit_model_rlm)
  
      # Find vertices
     vertices$Site_ID[i] <- sites[i]
     
    # get coeffs
    coefs_ma<- coef(parabola_fit_ma_rlm)
    c_ma <- unname(coefs_ma["(Intercept)"])
    b_ma <- unname(coefs_ma["Neon_SM"])
    a_ma <- unname(coefs_ma["I(Neon_SM^2)"])
    vertices$vertex_x_ma[i] <- (-b_ma/ (2*a_ma))
    vertices$vertex_y_ma[i] <- a_ma*(vertices$vertex_x_ma[i])^2 + b_ma*(vertices$vertex_x_ma[i]) + c_ma
    
     coefs_mq<- coef(parabola_fit_mq_rlm)
    c_mq <- unname(coefs_mq["(Intercept)"])
    b_mq <- unname(coefs_mq["Neon_SM"])
    a_mq <- unname(coefs_mq["I(Neon_SM^2)"])
    vertices$vertex_x_mq[i] <- (-b_mq/ (2*a_mq))
    vertices$vertex_y_mq[i] <- a_mq*(vertices$vertex_x_mq[i])^2 + b_mq*(vertices$vertex_x_mq[i]) + c_mq
    
     coefs_model<- coef(parabola_fit_model_rlm)
    c_model <- unname(coefs_model["(Intercept)"])
    b_model <- unname(coefs_model["Model_SM"])
    a_model <- unname(coefs_model["I(Model_SM^2)"])
    vertices$vertex_x_model[i] <- (-b_model/ (2*a_model))
    vertices$vertex_y_model[i] <- a_model*(vertices$vertex_x_model[i])^2 + b_model*(vertices$vertex_x_model[i]) + c_model
    

    # Graph
    
    print(ggplot() +
  #    geom_point(data = df_temp_ma, aes(x = Neon_SM, y = Marshall_SR, color = "NEON Data: Marshall Method"), size = 1) +
 #     geom_point(data = df_temp_mq, aes(x = Neon_SM, y = Mill_Quirk_SR, color = "NEON Data: Mill Quirk Method"), size = 1) +
  #    geom_point(data = df_temp_model, aes(x = Model_SM, y = Forecast_SR, color = "Model Output"), size = 1)+
 #     geom_line(data = df_temp_model, aes(x = Model_SM, y = y_fit_loess_model, color = "Model Output Fit"), size = 1)+
  #    geom_line(data = df_temp_mq, aes(x =Neon_SM, y = y_fit_loess_mq, color = "NEON Data Fit: Mill Quirk Method"), size = 1) +
  #    geom_line(data = df_temp_ma, aes(x = Neon_SM, y = y_fit_loess_ma, color = "NEON Data Fit: Marshall Method"), size = 1) +
      geom_line(data = df_temp_model, aes(x = Model_SM, y = y_fit_parabola_rlm_model, color = "Model Parabolic Fit"), size = 1)+
      geom_line(data = df_temp_mq, aes(x = Neon_SM, y = y_fit_parabola_rlm_mq, color = "NEON Parabolic Fit: Mill Quirk Method"), size = 1) +
      geom_line(data = df_temp_ma, aes(x = Neon_SM, y = y_fit_parabola_rlm_ma, color = "NEON Parabolic Fit: Marshall Method"), size = 1) +
      scale_color_manual(values = c("Model Output Fit" = "orange", "Model Output" = "black", "NEON Data: Mill Quirk Method" = "darkgrey", "NEON Data Fit: Mill Quirk Method" = "lightgreen", "NEON Data Fit: Marshall Method" = "darkgreen", "NEON Data: Marshall Method" = "lightgrey", "Model Parabolic Fit" = "orangered", "NEON Parabolic Fit: Mill Quirk Method" = "lightblue", "NEON Parabolic Fit: Marshall Method" = "darkblue")) +
      labs(title = sites[i],
           x = "Soil Moisture",
           y = "Soil Respiration",
           color = "Dataset"))

  
  
 }
   # combine ma and mq data frames
temp_df <- merge( df_temp_ma,  df_temp_mq, by = c("Time", "Site_ID", "Neon_SM"))
  
  # combine with model
  temp_df_final <- merge(temp_df,  df_temp_model, by = c("Time", "Site_ID"))

  SM_fits[[i]] <- temp_df_final
}
```

Graph Parabolic Fit for NEON & Model at each site 
```{r}


for (i in 1:length(sites)){
  temp_data <- filter(soil_table, Site_ID == sites[i])
  
  temp_data <- temp_data[order(temp_data$Neon_SM)]
  
  df_temp_ma <- data.frame(x = temp_data$Neon_SM,
                   y = temp_data$Marshall_SR)
  df_temp_ma <- df_temp_ma %>%
           filter(if_all(where(is.numeric), is.finite))
  
  df_temp_mq <- data.frame(x = temp_data$Neon_SM,
                   y = temp_data$Mill_Quirk_SR)
  df_temp_mq <- df_temp_mq %>%
           filter(if_all(where(is.numeric), is.finite))
  
  df_temp_model <- data.frame(x = temp_data$Model_SM,
                   y = temp_data$Forecast_SR)
  df_temp_model <- df_temp_model %>%
           filter(if_all(where(is.numeric), is.finite))
  
  if (is.na(df_temp_mq$x[1]) == FALSE){

  
      # Parabola fit 
    parabola_fit_ma_rlm <- rlm(y ~ x + I(x^2), data = df_temp_ma, na.action = na.exclude)
    df_temp_ma$y_fit_parabola_rlm <- predict(parabola_fit_ma_rlm)

     parabola_fit_mq_rlm <- rlm(y ~ x + I(x^2), data = df_temp_mq, na.action = na.exclude)
     df_temp_mq$y_fit_parabola_rlm <- predict(parabola_fit_mq_rlm)
     
     parabola_fit_model_rlm <- rlm(y ~ x + I(x^2), data = df_temp_model, na.action = na.exclude)
     df_temp_model$y_fit_parabola_rlm <- predict(parabola_fit_model_rlm)
    
# Graph
    print(ggplot() +
     # geom_point(data = df_temp_ma, aes(x = x, y = y, color = "NEON Data: Marshall Method"), size = 1) +
    #  geom_point(data = df_temp_mq, aes(x = x, y = y, color = "NEON Data: Mill Quirk Method"), size = 1) +
    #  geom_point(data = df_temp_model, aes(x = x, y = y, color = "Model Output"), size = 1)+
      geom_line(data = df_temp_model, aes(x = x, y = y_fit_parabola_rlm, color = "Model Parabolic Fit"), size = 1)+
      geom_line(data = df_temp_mq, aes(x = x, y = y_fit_parabola_rlm, color = "NEON Parabolic Fit: Mill Quirk Method"), size = 1) +
      geom_line(data = df_temp_ma, aes(x = x, y = y_fit_parabola_rlm, color = "NEON Parabolic Fit: Marshall Method"), size = 1) +
      scale_color_manual(values = c("Model Output Fit" = "orange", "Model Output" = "black", "NEON Data: Mill Quirk Method" = "darkgrey", "NEON Data Fit: Mill Quirk Method" = "lightgreen", "NEON Data Fit: Marshall Method" = "darkgreen", "NEON Data: Marshall Method" = "lightgrey", "Model Parabolic Fit" = "orangered", "NEON Parabolic Fit: Mill Quirk Method" = "lightblue", "NEON Parabolic Fit: Marshall Method" = "darkblue")) +
      labs(title = sites[i],
           x = "Soil Moisture",
           y = "Soil Respiration",
           color = "Dataset"))
  }

}
```


Soil respiration vs soil temperature, parabolic fit -- With reference point at vertex
```{r}

for (i in 1:length(sites)){
  
   # Get data for site
  temp_data <- SM_fits[[i]]
  
  if (is.na(vertices$vertex_x_ma[i])== FALSE){
    # Subtract vertex points from every point so that vertex is at zero
    temp_data$x_ma_ref <- temp_data$Neon_SM - vertices$vertex_x_ma[i]
    temp_data$y_fit_parabola_ma_ref_vertex <- temp_data$y_fit_parabola_rlm_ma - vertices$vertex_y_ma[i]
    
    temp_data$x_mq_ref <- temp_data$Neon_SM - vertices$vertex_x_mq[i]
    temp_data$y_fit_parabola_mq_ref_vertex <- temp_data$y_fit_parabola_rlm_mq - vertices$vertex_y_mq[i]
    
    temp_data$x_model_ref <- temp_data$Model_SM - vertices$vertex_x_model[i]
    temp_data$y_fit_parabola_model_ref_vertex <- temp_data$y_fit_parabola_rlm_model - vertices$vertex_y_model[i]
    
     print(ggplot() +
      geom_line(data = temp_data, aes(x = x_model_ref, y = y_fit_parabola_model_ref_vertex, color = "Model Parabolic Fit"), size = 1)+
      geom_line(data = temp_data, aes(x = x_mq_ref, y = y_fit_parabola_mq_ref_vertex, color = "NEON Parabolic Fit: Mill Quirk Method"), size = 1) +
      geom_line(data = temp_data, aes(x = x_ma_ref, y = y_fit_parabola_ma_ref_vertex, color = "NEON Parabolic Fit: Marshall Method"), size = 1) +
      scale_color_manual(values = c("Model Output Fit" = "orange", "Model Output" = "black", "NEON Data: Mill Quirk Method" = "darkgrey", "NEON Data Fit: Mill Quirk Method" = "lightgreen", "NEON Data Fit: Marshall Method" = "darkgreen", "NEON Data: Marshall Method" = "lightgrey", "Model Parabolic Fit" = "orangered", "NEON Parabolic Fit: Mill Quirk Method" = "lightblue", "NEON Parabolic Fit: Marshall Method" = "darkblue")) +
      labs(title = sites[i],
           x = "Soil Moisture",
           y = "Soil Respiration",
           color = "Dataset"))
  
  }
  
}

  
```


Soil respiration vs soil temperature, sigmoidal and loess regression fits -- With reference points at SM = 0.5
```{r}

for (i in 1:length(sites)){

  
  temp_data <- SM_fits[[i]]
  
  # -- Find the value closest to a set number 
  
  # Set the reference point
  target_value <- 0.5 
  
 # Get the models again
  df_temp_ma <- data.frame(Time = temp_data$Time,
                     Site_ID = temp_data$Site_ID,
                     Neon_SM = temp_data$Neon_SM,
                   Marshall_SR = temp_data$Marshall_SR)
  df_temp_ma <- df_temp_ma %>%
           filter(if_all(where(is.numeric), is.finite))
  
  df_temp_mq <- data.frame(Time = temp_data$Time,
                     Site_ID = temp_data$Site_ID,
                     Neon_SM = temp_data$Neon_SM,
                   Mill_Quirk_SR = temp_data$Mill_Quirk_SR)
  df_temp_mq <- df_temp_mq %>%
           filter(if_all(where(is.numeric), is.finite))
  
  df_temp_model <- data.frame(Time = temp_data$Time,
                     Site_ID = temp_data$Site_ID,
                     Model_SM = temp_data$Model_SM,
                   Forecast_SR = temp_data$Forecast_SR)
  df_temp_model <- df_temp_model %>%
           filter(if_all(where(is.numeric), is.finite))
  
    # Parabola fit 
    parabola_fit_ma_rlm <- rlm(Marshall_SR ~ Neon_SM + I(Neon_SM^2), data = df_temp_ma, na.action = na.exclude)
    temp_data$parabola_fit_ma_rlm <- predict(parabola_fit_ma_rlm)
     
    parabola_fit_mq_rlm <- rlm(Mill_Quirk_SR ~ Neon_SM + I(Neon_SM^2), data = df_temp_mq, na.action = na.exclude)
       temp_data$parabola_fit_mq_rlm <- predict(parabola_fit_mq_rlm)
     
     parabola_fit_model_rlm <- rlm(Forecast_SR ~ Model_SM + I(Model_SM^2), data = df_temp_model, na.action = na.exclude)
       temp_data$parabola_fit_model_rlm <- predict(parabola_fit_model_rlm)
  
     # Normalizing! 
    # Find the max of each
    max_val_ma <- max(temp_data$parabola_fit_ma_rlm)
    temp_data$y_fit_parabola_ma_norm <- temp_data$parabola_fit_ma_rlm/max_val_ma
    max_val_mq <- max(temp_data$parabola_fit_mq_rlm)
    temp_data$y_fit_parabola_mq_norm <- temp_data$parabola_fit_mq_rlm/max_val_mq
    max_val_model <- max(temp_data$parabola_fit_model_rlm)
    temp_data$y_fit_parabola_model_norm <- temp_data$parabola_fit_model_rlm/max_val_model
  
  # Find what y is at each x = target
    y_ma <- predict(parabola_fit_ma_rlm, data.frame(Neon_SM = target_value))/max_val_ma
    y_mq <- predict(parabola_fit_mq_rlm, data.frame(Neon_SM = target_value))/max_val_mq
    y_model <- predict(parabola_fit_model_rlm, data.frame(Model_SM = target_value))/max_val_model
     
  # -- Shift the graph down so that SR at 10 is 0 (for parabola fits)
  # if sig fits exist
  temp_data$y_fit_parabola_ma_ref <- temp_data$y_fit_parabola_ma_norm - y_ma
  temp_data$y_fit_parabola_mq_ref <- temp_data$y_fit_parabola_mq_norm - y_mq
  temp_data$y_fit_parabola_model_ref <- temp_data$y_fit_parabola_model_norm - y_model

  # graph
  pp <- ggplot() +
      geom_line(data = temp_data, aes(x = Neon_SM, y = y_fit_parabola_mq_ref, color = "Millington Quirk Sigmoidal Fit"), size = 1) +
      geom_line(data = temp_data, aes(x = Neon_SM, y = y_fit_parabola_ma_ref, color = "Marshall Sigmoidal Fit"), size = 1) +
      geom_line(data = temp_data, aes(x = Model_SM, y = y_fit_parabola_model_ref, color = "Model Sigmoidal Fit"), size = 1)  +  scale_color_manual(values = c("Model Loess Fit" = "orange", "Model Sigmoidal Fit" = "orangered", "Marshall Loess Fit" = "darkgreen", "Marshall Sigmoidal Fit" = "darkblue", "Millington Quirk Loess Fit" = "lightgreen", "Millington Quirk Sigmoidal Fit" = "lightblue")) +
      labs(title = sites[i], x = "Soil Temp", y = "Soil Respiration", color = "Dataset")
  
    # print the plot 
    print (pp)
    # Reassign temp fits so that it has reference data
    SM_fits[[i]] <- temp_data
}



```

Scoring with reference at SM = 0.5 (Normalized)
```{r}
# Make a table for score

sm_parabola_fit_scores <- data.frame(Site_ID = character(length(sites)),
                              MSD_ma = numeric(length(sites)),
                              MSD_mq = numeric(length(sites)),
                              CC_ma = numeric(length(sites)),
                              CC_mq = numeric(length(sites)),
                              Bias_ma = numeric (length(sites)),
                              Bias_mq = numeric (length(sites)),
                              Input_overlap = numeric (length(sites)))

# Loop through each site
for (i in 1:length(sites)){

  temp_data <- SM_fits[[i]]
  sm_parabola_fit_scores$Site_ID[i] <- sites[i]

  min_x <- max(min(temp_data$Neon_SM), min(temp_data$Model_SM));
  max_x <- min(max(temp_data$Neon_SM), max(temp_data$Model_SM));
  
  # Find input percent overlap
  percent_overlap <- ((max_x - min_x)/(max(max(temp_data$Neon_SM), max(temp_data$Model_SM))-min(min(temp_data$Neon_SM), min(temp_data$Model_SM)))) * 100
  
  sm_parabola_fit_scores$Input_overlap[i] <- round(percent_overlap)

    if ((min_x < max_x) == TRUE){
  
      
        # choose 10 points distributed across the range
       overlap_neon <- data.frame(Neon_SM = seq(from = min_x, to = max_x, length.out = 50))
       overlap_model <- data.frame(Model_SM = seq(from = min_x, to = max_x, length.out = 50))
       
       # Find the values at those points
       # Need to find function of fit again (because these points do not exist in data and don't want to use closest point)
       
       # Get the models again
    df_temp_ma <- data.frame(Time = temp_data$Time,
                       Site_ID = temp_data$Site_ID,
                       Neon_SM = temp_data$Neon_SM,
                     Marshall_SR = temp_data$Marshall_SR)
    df_temp_ma <- df_temp_ma %>%
             filter(if_all(where(is.numeric), is.finite))
    
    df_temp_mq <- data.frame(Time = temp_data$Time,
                       Site_ID = temp_data$Site_ID,
                       Neon_SM = temp_data$Neon_SM,
                     Mill_Quirk_SR = temp_data$Mill_Quirk_SR)
    df_temp_mq <- df_temp_mq %>%
             filter(if_all(where(is.numeric), is.finite))
    
    df_temp_model <- data.frame(Time = temp_data$Time,
                       Site_ID = temp_data$Site_ID,
                       Model_SM = temp_data$Model_SM,
                     Forecast_SR = temp_data$Forecast_SR)
    df_temp_model <- df_temp_model %>%
             filter(if_all(where(is.numeric), is.finite))
  
    # Parabola fit 
    parabola_fit_ma_rlm <- rlm(Marshall_SR ~ Neon_SM + I(Neon_SM^2), data = df_temp_ma, na.action = na.exclude)
    parabola_fit_mq_rlm <- rlm(Mill_Quirk_SR ~ Neon_SM + I(Neon_SM^2), data = df_temp_mq, na.action = na.exclude)
    parabola_fit_model_rlm <- rlm(Forecast_SR ~ Model_SM + I(Model_SM^2), data = df_temp_model, na.action = na.exclude)
    
    # Set reference SM
    target_value = 0.5
  
    # Find value at target so that this is based off reference point
    y_ma <- predict(parabola_fit_ma_rlm, data.frame(Neon_SM = target_value))[1]
    y_mq <- predict(parabola_fit_mq_rlm, data.frame(Neon_SM = target_value))[1]
    y_model <- predict(parabola_fit_model_rlm, data.frame(Model_SM = target_value))[1]
    
    # Add bias to table
    sm_parabola_fit_scores$Bias_ma[i] <- y_model - y_ma
    sm_parabola_fit_scores$Bias_mq[i] <- y_model - y_mq
    
     # find max values to normalize
       max_val_ma <- max(temp_data$parabola_fit_ma_rlm)
       max_val_mq <- max(temp_data$parabola_fit_mq_rlm)
       max_val_model <- max(temp_data$parabola_fit_model_rlm)
    
      # use function to find values at points
       actual <- (predict(parabola_fit_model_rlm, overlap_model)[1:50] - y_model)/max_val_model
       predicted_ma <- (predict(parabola_fit_ma_rlm, overlap_neon)[1:50] - y_ma)/max_val_ma
       predicted_mq <- (predict(parabola_fit_mq_rlm, overlap_neon)[1:50] - y_mq)/max_val_mq
       
      # Find mean((actual-predicted)^2)
       msd_ma <-  mean((actual-predicted_ma)^2)
       msd_mq <-  mean((actual-predicted_mq)^2)
       
       sm_parabola_fit_scores$MSD_ma[i] <- round(msd_ma, digits = 2)
       sm_parabola_fit_scores$MSD_mq[i] <- round(msd_mq, digits = 2)

       
      # Correlation Coeffs
       sm_parabola_fit_scores$CC_ma[i] <- round(cor(actual, predicted_ma), digits = 2);
       sm_parabola_fit_scores$CC_mq[i] <- round(cor(actual, predicted_mq), digits = 2);
    
  
  # Put into table to compare
    }else{
      # If range doesn't over lap, add "NA" to table
      sm_parabola_fit_scores$MSD_ma[i] <- NA
      sm_parabola_fit_scores$MSD_mq[i] <- NA
    }
    
}


```


Scoring with reference at vertex
```{r}

# Make a table for score
sm_parabola_fit_scores <- data.frame(Site_ID = character(length(sites)),
                              MSD_ma = numeric(length(sites)),
                              MSD_mq = numeric(length(sites)),
                              Shape_score_ma = numeric(length(sites)),
                              Shape_score_mq = numeric(length(sites)),
                              Input_score_ma = numeric(length(sites)),
                              Input_score_mq = numeric(length(sites)))

# Loop through each site
for (i in 1:length(sites)){

  temp_data <- SM_fits[[i]]
  sm_parabola_fit_scores$Site_ID[i] <- sites[i]

  min_x <- max(min(temp_data$Neon_SM), min(temp_data$Model_SM));
  max_x <- min(max(temp_data$Neon_SM), max(temp_data$Model_SM));

    if ((min_x < max_x) == TRUE){
      
        # Choose 10 points distributed across the range
       overlap_neon <- data.frame(Neon_SM = seq(from = min_x, to = max_x, length.out = 50))
       overlap_model <- data.frame(Model_SM = seq(from = min_x, to = max_x, length.out = 50))
       
       # Find the values at those points
       # Get the models again
    df_temp_ma <- data.frame(Time = temp_data$Time,
                       Site_ID = temp_data$Site_ID,
                       Neon_SM = temp_data$Neon_SM,
                     Marshall_SR = temp_data$Marshall_SR)
    df_temp_ma <- df_temp_ma %>%
             filter(if_all(where(is.numeric), is.finite))
    
    df_temp_mq <- data.frame(Time = temp_data$Time,
                       Site_ID = temp_data$Site_ID,
                       Neon_SM = temp_data$Neon_SM,
                     Mill_Quirk_SR = temp_data$Mill_Quirk_SR)
    df_temp_mq <- df_temp_mq %>%
             filter(if_all(where(is.numeric), is.finite))
    
    df_temp_model <- data.frame(Time = temp_data$Time,
                       Site_ID = temp_data$Site_ID,
                       Model_SM = temp_data$Model_SM,
                     Forecast_SR = temp_data$Forecast_SR)
    df_temp_model <- df_temp_model %>%
             filter(if_all(where(is.numeric), is.finite))
  
    # Parabola fit 
    parabola_fit_ma_rlm <- rlm(Marshall_SR ~ Neon_SM + I(Neon_SM^2), data = df_temp_ma, na.action = na.exclude)
    parabola_fit_mq_rlm <- rlm(Mill_Quirk_SR ~ Neon_SM + I(Neon_SM^2), data = df_temp_mq, na.action = na.exclude)
    parabola_fit_model_rlm <- rlm(Forecast_SR ~ Model_SM + I(Model_SM^2), data = df_temp_model, na.action = na.exclude)
  
      # Use function to find values at points
       actual <- predict(parabola_fit_model_rlm, overlap_model)[1:50]
       predicted_ma <- predict(parabola_fit_ma_rlm, overlap_neon)[1:50]
       predicted_mq <- predict(parabola_fit_mq_rlm, overlap_neon)[1:50]
       
      # Find mean((actual-predicted)^2)
       msd_ma <-  mean((actual-predicted_ma)^2)
       msd_mq <-  mean((actual-predicted_mq)^2)
       sm_parabola_fit_scores$MSD_ma[i] <- msd_ma
       sm_parabola_fit_scores$MSD_mq[i] <- msd_mq
       
      sm_parabola_fit_scores$Shape_score_ma[i] <- round(msd_ma*10^16)
      sm_parabola_fit_scores$Shape_score_mq[i] <- round(msd_mq*10^16)
  
  # Put into table to compare
    }else{
      # If range doesn't over lap, add "NA" to table
      sm_parabola_fit_scores$MSD_ma[i] <- "No overlap"
      sm_parabola_fit_scores$MSD_mq[i] <- "No overlap"
       
      sm_parabola_fit_scores$Shape_score_ma[i] <- "No overlap"
      sm_parabola_fit_scores$Shape_Shape_score_mq[i] <- "No overlap"
    }
    
}

```

Compare scores to other factors

Precipitiation
```{r}
# Find average annual precipitation 

load("C:/Users/chane/OneDrive/Documents/ef_lab/R_files/precip_data.RData")

precip_data_ave <- precip_data_day %>%
  group_by(Site_ID) %>%
  summarize(across(c(precipBulk), mean, na.rm = TRUE), .groups = 'drop')

sm_scores <- merge(sm_parabola_fit_scores, precip_data_ave, by = "Site_ID")

# Plot score vs average precipitation

ggplot()+
  geom_point(data = sm_scores, aes(x = precipBulk, y = MSD_ma, color = "Marshall Mean Square Distance"))+
  geom_point(data = sm_scores, aes(x = precipBulk, y =  MSD_mq, color = "Mill Quirk Mean Square Distance"))+
  geom_smooth(data = sm_scores, aes(x = precipBulk, y = MSD_ma, color = "red"), method = "lm", se = FALSE)+
  geom_smooth(data = sm_scores, aes (x = precipBulk, y =  MSD_mq, color = "orange"), method = "lm", se = FALSE)+
  scale_color_manual(values = c("Marshall Mean Square Distance" = "darkgreen", "Mill Quirk Mean Square Distance" = "lightgreen"))+
  labs(title = "", y = "Mean Square Distance", x = "Average Annual Precipitation", color = "Dataset")

ggplot()+
  geom_point(data = sm_scores, aes(x = precipBulk, y = CC_ma, color = "Marshall Correlation Coefficient"))+
  geom_point(data = sm_scores, aes(x = precipBulk, y = CC_mq, color = "Mill Quirk Correlation Coefficient"))+
  geom_smooth(data = sm_scores, aes(x = precipBulk, y = CC_ma, color = "red"), method = "lm", se = FALSE)+
  geom_smooth(data = sm_scores, aes (x = precipBulk, y = CC_mq, color = "orange"), method = "lm", se = FALSE)+
  scale_color_manual(values = c("Marshall Correlation Coefficient" = "darkgreen", "Mill Quirk Correlation Coefficient" = "lightgreen"))+
  labs(title = "", y = "Correlation Coefficient", x = "Average Annual Precipitation", color = "Dataset")

ggplot()+
  geom_point(data = sm_scores, aes(x = precipBulk, y = Bias_ma, color = "Marshall"))+
  geom_point(data = sm_scores, aes(x = precipBulk, y = Bias_mq, color = "Mill Quirk"))+
  geom_smooth(data = sm_scores, aes(x = precipBulk, y = Bias_ma, color = "red"), method = "lm", se = FALSE)+
  geom_smooth(data = sm_scores, aes (x = precipBulk, y = Bias_mq, color = "orange"), method = "lm", se = FALSE)+
  scale_color_manual(values = c("Marshall" = "darkgreen", "Mill Quirk" = "lightgreen"))+
  labs(title = "", y = "Bias", x = "Average Annual Precipitation", color = "Dataset")

ggplot()+
  geom_point(data = sm_scores, aes(x = precipBulk, y = Input_overlap), color = "grey")+
  geom_smooth(data = sm_scores, aes(x = precipBulk, y = Input_overlap), color = "black", method = "lm", se = FALSE)+
  labs(title = "", y = "Input Overlap", x = "Average Annual Precipitation", color = "Dataset")



```





Compare to 13 ecological factors
```{r}
dat = terra::rast("/projectnb/dietzelab/dongchen/anchorSites/downscale/all_data_layers.tif")

# add columns to error_df
comparisons_table <- data.frame(Site_ID = character(length(sites)),
                    latitude = numeric(length(sites)),
                    longitude = numeric(length(sites)),
                    ysd =numeric(length(sites)),
                    abg =  numeric(length(sites)),
                    fia = numeric(length(sites)),
                    gedi = numeric(length(sites)),
                    twi = numeric(length(sites)),
                    tavg = numeric(length(sites)),
                    srad = numeric(length(sites)),
                    prec = numeric(length(sites)),
                    vapr = numeric(length(sites)),
                    ph = numeric(length(sites)),
                    n = numeric(length(sites)),
                    soc = numeric(length(sites)),
                    sand = numeric(length(sites)))

pt = numeric(1)

# Extract data for each from each Neon site coordinate
for (i in 1:length(sites)){
  site_num = which(Neonplots$field_site_id == sites[i])
  pt = terra::extract(dat,data.frame(lon=Neonplots$field_longitude[site_num], lat = Neonplots$field_latitude[site_num]))
  comparisons_table$Site_ID[i] <- Neonplots$field_site_id[site_num]
  comparisons_table$Latitude[i] <- Neonplots$field_latitude[site_num]
  comparisons_table$Longitude[i] <- Neonplots$field_longitude[site_num]
  
  comparisons_table$ysd[i] = pt$year_since_disturb
  comparisons_table$abg[i] <- pt$agb
  comparisons_table$fia[i] <- pt$fia
  comparisons_table$gedi[i] <- pt$gedi
  comparisons_table$twi[i] <- pt$twi
  comparisons_table$tavg[i] <- pt$tavg
  comparisons_table$srad[i] <- pt$srad
  comparisons_table$prec[i] <- pt$prec
  comparisons_table$vapr[i] <- pt$vapr
  comparisons_table$ph[i] <- pt$PH
  comparisons_table$n[i] <- pt$N
  comparisons_table$soc[i] <- pt$SOC
  comparisons_table$sand[i] <- pt$Sand
}



# Or Load from SCC
load("all_data_layers.Rdata")

sm_scores <- merge(sm_scores, comparisons_table, by = "Site_ID")

```


Latitude
```{r}
ggplot()+
  geom_point(data = sm_scores, aes(x = Latitude, y =  MSD_ma, color = "Marshall Mean Square Distance"))+
  geom_point(data = sm_scores, aes(x = Latitude, y = MSD_mq, color = "Mill Quirk Mean Square Distance"))+
  geom_smooth(data = sm_scores, aes(x = Latitude, y = MSD_ma, color = "red"), method = "lm", se = FALSE)+
  geom_smooth(data = sm_scores, aes (x = Latitude, y = MSD_mq, color = "orange"), method = "lm", se = FALSE)+
  scale_color_manual(values = c("Marshall Mean Square Distance" = "darkgreen", "Mill Quirk Mean Square Distance" = "lightgreen"))+
  labs(title = "", y = "Mean Square Distance", x = "Latitude", color = "Dataset")

ggplot()+
  geom_point(data = sm_scores, aes(x = Latitude, y = CC_ma, color = "Marshall Correlation Coefficient"))+
  geom_point(data = sm_scores, aes(x = Latitude, y = CC_mq, color = "Mill Quirk Correlation Coefficient"))+
  geom_smooth(data = sm_scores, aes(x = Latitude, y = CC_ma, color = "red"), method = "lm", se = FALSE)+
  geom_smooth(data = sm_scores, aes (x = Latitude, y =  CC_mq, color = "orange"), method = "lm", se = FALSE)+
  scale_color_manual(values = c("Marshall Correlation Coefficient" = "darkgreen", "Mill Quirk Correlation Coefficient" = "lightgreen"))+
  labs(title = "", y = "Correlation Coefficient", x = "Latitude", color = "Dataset")

ggplot()+
  geom_point(data = sm_scores, aes(x = Latitude, y = Bias_ma, color = "Marshall"))+
  geom_point(data = sm_scores, aes(x = Latitude, y = Bias_mq, color = "Mill Quirk"))+
  geom_smooth(data = sm_scores, aes(x = Latitude, y = Bias_ma, color = "red"), method = "lm", se = FALSE)+
  geom_smooth(data = sm_scores, aes (x = Latitude, y = Bias_mq, color = "orange"), method = "lm", se = FALSE)+
  scale_color_manual(values = c("Marshall" = "darkgreen", "Mill Quirk" = "lightgreen"))+
  labs(title = "", y = "Bias", x = "Latitude", color = "Dataset")

ggplot()+
  geom_point(data = sm_scores, aes(x = Latitude, y = Input_overlap), color = "grey")+
  geom_smooth(data = sm_scores, aes(x = Latitude, y = Input_overlap), color = "black", method = "lm", se = FALSE)+
  labs(title = "", y = "Input Overlap", x = "Latitude", color = "Dataset")


```

```{r}
for (i in (length(sm_scores) - 12):length(sm_scores)){
  # fit a line
  sm_scores$temp = sm_scores[[i]]
  
  # graph against scores
   print(ggplot()+
  geom_point(data = sm_scores, aes(x = temp, y = MSD_ma, color = "Marshall Correlation Coefficient"))+
  geom_point(data = sm_scores, aes(x = temp, y = MSD_mq, color = "Mill Quirk Correlation Coefficient"))+
  geom_smooth(data = sm_scores, aes(x = temp, y = MSD_ma, color = "Marshall Line Fit"), method = "lm", se = FALSE)+
  geom_smooth(data = sm_scores, aes (x = temp, y =  MSD_mq, color = "Mill Quirk Line Fit"), method = "lm", se = FALSE)+
  scale_color_manual(values = c("Marshall Correlation Coefficient" = "darkgreen", "Mill Quirk Correlation Coefficient" = "lightgreen", "Marshall Line Fit" = "black", "Mill Quirk Line Fit" = "darkgray"))+
  labs(title = "", y = "MSD", x = colnames(sm_scores[(i)]), color = "Dataset"))


 print(ggplot()+
  geom_point(data = sm_scores, aes(x = temp, y = CC_ma, color = "Marshall Correlation Coefficient"))+
  geom_point(data = sm_scores, aes(x = temp, y = CC_mq, color = "Mill Quirk Correlation Coefficient"))+
  geom_smooth(data = sm_scores, aes(x = temp, y = CC_ma, color = "Marshall Line Fit"), method = "lm", se = FALSE)+
  geom_smooth(data = sm_scores, aes (x = temp, y =  CC_mq, color = "Mill Quirk Line Fit"), method = "lm", se = FALSE)+
  scale_color_manual(values = c("Marshall Correlation Coefficient" = "darkgreen", "Mill Quirk Correlation Coefficient" = "lightgreen", "Marshall Line Fit" = "black", "Mill Quirk Line Fit" = "darkgray"))+
  labs(title = "", y = "Correlation Coefficient", x = colnames(sm_scores[(i)]), color = "Dataset"))
 
 print(ggplot()+
  geom_point(data = sm_scores, aes(x = temp, y =  Bias_ma, color = "Marshall Correlation Coefficient"))+
  geom_point(data = sm_scores, aes(x = temp, y =  Bias_mq, color = "Mill Quirk Correlation Coefficient"))+
  geom_smooth(data = sm_scores, aes(x = temp, y =  Bias_ma, color = "Marshall Line Fit"), method = "lm", se = FALSE)+
  geom_smooth(data = sm_scores, aes (x = temp, y =   Bias_mq, color = "Mill Quirk Line Fit"), method = "lm", se = FALSE)+
  scale_color_manual(values = c("Marshall Correlation Coefficient" = "darkgreen", "Mill Quirk Correlation Coefficient" = "lightgreen", "Marshall Line Fit" = "black", "Mill Quirk Line Fit" = "darkgray"))+
  labs(title = "", y = "Bias", x = colnames(sm_scores[(i)]), color = "Dataset"))
 
 print(ggplot()+
  geom_point(data = sm_scores, aes(x = temp, y = Input_overlap), color = "grey")+
  geom_smooth(data = sm_scores, aes(x = temp, y = Input_overlap), color = "black", method = "lm", se = FALSE)+
  labs(title = "", y = "Input Overlap", x = colnames(sm_scores[(i)]), color = "Dataset"))

 
}
```

