---
title: "soilrespiration_validation"
author: "chaney"
date: "2025-11-24"
output: html_document
---

All validation for soil respiration, not including validations that involve moisture and temperature. 

```{r setup, include=FALSE}
# Load libraries

library(DBI)
library(dplyr)
library(geosphere)
library(sp)
library(ggplot2)
library(maps)
library(knitr)
library(tidyr)
library(neonUtilities)
library(maps)
library(tidyverse)
library(neonSoilFlux)
library(parallel)
library(ncdf4)
library(data.table)
library(purrr)
library(data.table)
library(parallel)
library(lubridate)
library(httr)
library(sp)
library(neonSoilFlux)
library(stringr)
library(scales)


```

First, find model plots for each NEON plot
```{r}

# get the coordinates of 6400 sites in Dongchen's output
Neonplots <- read.csv("/usr4/ugrad/chaney/R/3_25/Neon_sites_terrestrial.csv") 

# Add a column for site numbers in Site_Info (site_info has coordinates of each location in df)
Site_Info <- readRDS("/usr4/ugrad/chaney/R/3_25/site.locs.rds")
Site_Info$Site_Number <- seq(1, nrow(Site_Info)) 

# Extract latitude and longitude from Neonplots and Site_Info tables
neon_coords <- Neonplots[ ,2:3] %>%
  select(Latitude = field_latitude, Longitude = field_longitude)
Site_Info_coords <- Site_Info %>%
  select(Site_Number, lat, lon)

# Put Site_Info latitude/longitude in a matrix
Site_Coordinates <- as.matrix(Site_Info[, c("lat", "lon")])

# Same for Neonplots
Neonplots_Coordinates <- as.matrix(Neonplots[, 2:3])

# make list to store matched list
matched <- data.frame(id = numeric(nrow(Neonplots)), # id in forecast
                      site_id = numeric(nrow(Neonplots))) # NEON id

# Loop through each row in Neonplots and calculate each plot's closest site number
for (i in 1:nrow(Neonplots)) {
  # Extract the coordinates for the current row in Neonplots
  plot_coords <- Neonplots_Coordinates[i, ]
  
  # Compute distances from this plot to all site coordinates (longlat means its in km)
  distances <- spDistsN1(Site_Coordinates, plot_coords, longlat = TRUE)
  
  # Find the index of the minimum distance for closest site
  closest_site_index <- which.min(distances)
  
  # Assign the corresponding Site_Number from Site_Info to the current row in Neonplots
  matched$id[i] <- Site_Info$Site_Number[closest_site_index]
  matched$site_id[i] <- Neonplots$field_site_id[i]
}

```


Get soil respiration output from SDA model
```{r}

source("model_processing_neonsites.R")

# Set outdir
outdir <- "/projectnb/dietzelab/dongchen/anchorSites/NA_runs/SDA_8k_site/NA_SDA_no_Debias";

final_daily_sr_data <- model_processing_neonsites(matched, 2012, 2025, outdir, "SoilResp", var2 = NULL)

colnames(final_daily_sr_data)[colnames(final_daily_sr_data) == "Output1"] <- "Model_SR"

```


Get NEON Soil respiration from John's data in the SCC
```{r}
# Function to get soil fluxes for specifed years for 1 site

soil_flux <- function(site_number, flux_type){

  years = 2012:2025
  site = matched$site_id[site_number]
  
  # create list to store data
  all_days <- list() 
  count <- 1 

  for (year in years){
    for (i in 01:12){
      if (i < 10){
        file_name <- paste0("/projectnb/dietzelab/jzobitz/02-NEON-sites/flux-results/out-flux-", site, "-",year,"-0",i,".Rda")
      }else{
        file_name <- paste0("/projectnb/dietzelab/jzobitz/02-NEON-sites/flux-results/out-flux-", site, "-",year,"-",i,".Rda")
      }
      if (file.exists(file_name)== TRUE){
        load(file_name)
        
        # sort by date so I can average over all horizontal positions
        out_fluxes <- out_fluxes[order(out_fluxes$startDateTime),]
        
        #  Exatract all flux values into a vector 
        flux_vec <- sapply(out_fluxes[[3]], function(x) x$flux[flux_type])
        
        # Fine the number of data points per day
        dppd = 48 * length(unique(out_fluxes$horizontalPosition))
        
        n_days <- length(flux_vec) / (dppd)
        
        # Create day index (every 240 rows = 1 day)
        day_idx <- ceiling(seq_along(flux_vec) / dppd)
        
        # Count rows in each day
        day_sizes <- table(day_idx)

        # Identify complete days only
        complete_days <- as.numeric(names(day_sizes[day_sizes == dppd]))

        # Filter vectors to ONLY complete days
        keep <- day_idx %in% complete_days
        
        flux_vec_f <- flux_vec[keep]

        # Recompute number of complete days
        n_days <- length(flux_vec_f) / dppd
          
        # Turn vector into column where column = day
        flux_mat <- matrix(flux_vec_f, nrow = dppd, ncol = n_days)
        
        # Find daily means
        daily_means <- colMeans(flux_mat, na.rm = TRUE)
        
        # Find times
        day_times <- out_fluxes$startDateTime[seq(1, length(flux_vec_f), by = dppd)]
        
        # Add to table
        summary_table <- data.frame(
          Site_ID = site,
          Time = day_times,
          sr_neon = daily_means
        )
        
        # Add table to list
      all_days[[count]] <- summary_table
      count <- count + 1
      }
    }
  }
  # turn list into table
  summary <- rbindlist(all_days)
  
return(list(sr = summary))
}


# Apply parallel calculation for each site
n = 1:nrow(matched)
num_cores <- detectCores() - 1

# using Marshall method at top of soil (000)
results_j <- mclapply(n, soil_flux, mc.cores = num_cores, flux_type = 4)

# Combine results from all ensmebles
sr_marshall <- rbindlist(lapply(results_j, `[[`, "sr"))

# delete rows with all zeros
sr_marshall_filtered <- filter(sr_marshall, sr_neon != 0)

# using Millington- Quirk method at top of soil (000)
results_j_2 <- mclapply(n, soil_flux, mc.cores = num_cores, flux_type = 8)

# Combine results from all ensembles
sr_m_q <- rbindlist(lapply(results_j_2, `[[`, "sr"))

# delete rows where the soil respiration is 0
sr_m_q_filtered <-filter(sr_m_q, sr_neon != 0)


# combine the two methods
daily_mean_data_neon <- merge(sr_marshall_filtered, sr_m_q_filtered, by = c("Time", "Site_ID"))

# rename rows
names(daily_mean_data_neon)[names(daily_mean_data_neon) == "sr_neon.x"] <- "Marshall_SR"
names(daily_mean_data_neon)[names(daily_mean_data_neon) == "sr_neon.y"] <- "Mill_Quirk_SR"

# convert Time to date class
final_daily_sr_data$Time <- as.Date(final_daily_sr_data$Time)

# combine data frame into final_daily_sr_data
daily_sr <- merge(final_daily_sr_data, daily_mean_data_neon, by = c("Time","Site_ID"))
names(daily_sr)[names(daily_sr) == "Final_SR"] <- "Forecast_SR"

# Convert NEON data to kg/mol ^2 *s (currently in umol/m^2 * s)
daily_sr$Marshall_SR = daily_sr$Marshall_SR*10^-6*12.011*10^-3
daily_sr$Mill_Quirk_SR = daily_sr$Mill_Quirk_SR*10^-6*12.011*10^-3
```

Combine NEON & Model into one table
```{r}
# combine the two methods
daily_mean_data_neon <- merge(sr_marshall_filtered, sr_m_q_filtered, by = c("Time", "Site_id"))

# rename rows
names(daily_mean_data_neon)[names(daily_mean_data_neon) == "sr_neon.x"] <- "Marshall_SR"
names(daily_mean_data_neon)[names(daily_mean_data_neon) == "sr_neon.y"] <- "Mill_Quirk_SR"
names(daily_mean_data_neon)[names(daily_mean_data_neon) == "Site_id"] <- "Site_ID"

# convert Time to date class
final_daily_sr_data$Time <- as.Date(final_daily_sr_data$Time)
daily_mean_data_neon$Time <- as.Date(daily_mean_data_neon$Time)

# Delete var2
final_daily_sr_data$Output2 <- NULL

# combine data frame into final_daily_sr_data
daily_sr <- merge(final_daily_sr_data, daily_mean_data_neon, by = c("Time","Site_ID"))

# Convert NEON data to kg/mol ^2 *s (currently in umol/m^2 * s)
daily_sr$Marshall_SR = daily_sr$Marshall_SR*10^-6*12.011*10^-3
daily_sr$Mill_Quirk_SR = daily_sr$Mill_Quirk_SR*10^-6*12.011*10^-3

```

Graph Predicted-Observed Plots for each site
```{r}

for (i in 1:length(sites)){
  temp_data <- filter(daily_sr, Site_ID == sites[i])
  
  print(ggplot() + 
        geom_point(data = temp_data, aes(x = Forecast_SR, y = Marshall_SR, color = "Marshall"))+ 
     #   geom_line(data = temp_data, aes(x = Forecast_SR, y = Marshall_SR, color = "Marshall Fit"), size = 1) + 
        geom_point(data = temp_data, aes(x = Forecast_SR, y = Mill_Quirk_SR, color = "Millington-Quirk")) + 
    #    geom_line(data = temp_data, aes(x = Forecast_SR, y = Mill_Quirk_SR, color = "Millington-Quirk Fit"), size = 1) + 
        geom_abline(slope = 1, intercept = 0, color = "red", size = 2)+
        scale_color_manual(values = c("Marshall" = "darkgreen", "Millington-Quirk" = "lightgreen", "Marshall Fit" = "darkgreen", "Mill Quirk Fit" = "lightgreen")) +
        labs(title = sites[i], x = "Predicted (kg/m^2*s)", y = "Observed (kg/m^2*s)", color = "Dataset"))

}
```


Calculate Summary Statistics
```{r}
# Calculate error
daily_sr$Marsh_error <- daily_sr$Forecast_SR - daily_sr$Marshall_SR
daily_sr$MQ_error <- daily_sr$Forecast_SR - daily_sr$Mill_Quirk_SR

# keep a permanent daily_sr
daily_sr_perm <- daily_sr

# make table for summary statistics
summary_stat <- data.frame(Site_ID = character(length(sites)),
                           Marsh_corr_coef = numeric(length(sites)),
                           MQ_corr_coef = numeric(length(sites)),
                           Marsh_rms_error = numeric(length(sites)),
                           MQ_rms_error = numeric(length(sites)))


# do summary stats
sites <- unique(daily_sr$Site_ID)

for (i in 1:length(sites)){
  # filter for just one site
  temp_data <- filter(daily_sr, Site_ID == sites[i])
  
  # remove na values
  temp_data <- na.omit(temp_data)
  
  # get rid of first row (so i can remove infinite)
  temp_data <- subset(temp_data, select = -c(Site_ID, Time))
  
  # remove inf values
  temp_data <- temp_data[apply(temp_data, 1, function(row) all(is.finite(row))), ]
  
  # add site name
  summary_stat$Site_ID[i] <- sites[i]
  
  # find correlation coeffs
  summary_stat$Marsh_corr_coef[i] <- cor(temp_data$Marshall_SR, temp_data$Forecast_SR)
  summary_stat$MQ_corr_coef[i] <- cor(temp_data$Mill_Quirk_SR, temp_data$Forecast_SR)
  
  # find RMS errors
  summary_stat$Marsh_rms_error[i] <- sqrt(mean((temp_data$Marshall_SR - temp_data$Forecast_SR)^2))
  summary_stat$MQ_rms_error[i] <- sqrt(mean((temp_data$Mill_Quirk_SR - temp_data$Forecast_SR)^2))
}


```

Add latitude and longitude to make maps!
```{r}
# Add site coordinates
summary_stat$Site_Lat <- numeric(nrow(summary_stat))
summary_stat$Site_Long <- numeric(nrow(summary_stat))

for (i in 1:length(summary_stat$Site_ID)){
  site_num = which(Neonplots$field_site_id == summary_stat$Site_ID[i])
  summary_stat$Site_Lat[i] <- Neonplots$field_latitude[site_num]
  summary_stat$Site_Long[i] <- Neonplots$field_longitude[site_num]
}

```

Maps of Correlation Coefficient
```{r}
# map of correlation coefficient
print(ggplot()+
  geom_polygon(data = USA, aes(x= long, y = lat, group = group), fill="forestgreen", alpha = 0.3) + geom_point(data = summary_stat, aes(x= Site_Long, y = Site_Lat, size = Marsh_corr_coef, color = Marsh_corr_coef)) + geom_point(alpha = 0.7) + xlim(-170, -50) + scale_size_continuous(range = c(1,10)) + theme_void() + labs(title="Marshall Correlation Coefficient") + theme(plot.title = element_text(hjust = 0.5),))

print(ggplot()+
  geom_polygon(data = USA, aes(x= long, y = lat, group = group), fill="forestgreen", alpha = 0.3) + geom_point(data = summary_stat, aes(x= Site_Long, y = Site_Lat, size = MQ_corr_coef, color = MQ_corr_coef)) + geom_point(alpha = 0.7) + xlim(-170, -50) + scale_size_continuous(range = c(1,10)) + theme_void() + labs(title="MQ Correlation Coefficient") + theme(plot.title = element_text(hjust = 0.5),))
```

Maps of RMS error
```{r}
# map of RMS error
print(ggplot()+
  geom_polygon(data = USA, aes(x= long, y = lat, group = group), fill="forestgreen", alpha = 0.3) + geom_point(data = summary_stat, aes(x= Site_Long, y = Site_Lat, size = Marsh_rms_error, color = Marsh_rms_error)) + geom_point(alpha = 0.7) + xlim(-170, -50) + scale_size_continuous(range = c(1,10)) + theme_void() + labs(title="Marshall RMS error") + theme(plot.title = element_text(hjust = 0.5),))

print(ggplot()+
  geom_polygon(data = USA, aes(x= long, y = lat, group = group), fill="forestgreen", alpha = 0.3) + geom_point(data = summary_stat, aes(x= Site_Long, y = Site_Lat, size = MQ_rms_error, color = MQ_rms_error)) + geom_point(alpha = 0.7) + xlim(-170, -50) + scale_size_continuous(range = c(1,10)) + theme_void() + labs(title="MQ RMS error") + theme(plot.title = element_text(hjust = 0.5),))
```


Compare error to a variety of ecological factors 
```{r}
# get data on US 13 factors
dat = terra::rast("all_data_layers.tif")

# add columns to error_df
summary_stat$ysd <- numeric(length(sites))
summary_stat$abg<- numeric(length(sites))
summary_stat$fia <- numeric(length(sites))
summary_stat$gedi<- numeric(length(sites))
summary_stat$twi<- numeric(length(sites))
summary_stat$tavg <- numeric(length(sites))
summary_stat$srad <- numeric(length(sites))
summary_stat$prec <- numeric(length(sites))
summary_stat$vapr <- numeric(length(sites))
summary_stat$ph <- numeric(length(sites))
summary_stat$n <- numeric(length(sites))
summary_stat$soc <- numeric(length(sites))
summary_stat$sand <- numeric(length(sites))

pt = numeric(1)

# extract data for each from each Neon site coordinate
for (i in 1:nrow(summary_stat)){
  pt = terra::extract(dat,data.frame(lon=summary_stat$Site_Long[i], lat = summary_stat$Site_Lat[i]))
  summary_stat$ysd[i] <- pt$year_since_disturb
  summary_stat$abg[i] <- pt$agb
  summary_stat$fia[i] <- pt$fia
  summary_stat$gedi[i] <- pt$gedi
  summary_stat$twi[i] <- pt$twi
  summary_stat$tavg[i] <- pt$tavg
  summary_stat$srad[i] <- pt$srad
  summary_stat$prec[i] <- pt$prec
  summary_stat$vapr[i] <- pt$vapr
  summary_stat$ph[i] <- pt$PH
  summary_stat$n[i] <- pt$N
  summary_stat$soc[i] <- pt$SOC
  summary_stat$sand[i] <- pt$Sand
}

# use lm to fit a linear model to the data
fit_ysd = lm(summary_stat$ysd ~ summary_stat$Marsh_corr_coef)
fit_abg = lm(summary_stat$abg ~ summary_stat$Marsh_corr_coef)
fit_fia = lm(summary_stat$fia ~ summary_stat$Marsh_corr_coef)
fit_gedi = lm(summary_stat$gedi ~ summary_stat$Marsh_corr_coef)
fit_twi = lm(summary_stat$twi ~ summary_stat$Marsh_corr_coef)
fit_tavg = lm(summary_stat$tavg ~ summary_stat$Marsh_corr_coef)
fit_srad = lm(summary_stat$srad ~ summary_stat$Marsh_corr_coef)
fit_prec = lm(summary_stat$prec~ summary_stat$Marsh_corr_coef)
fit_vapr = lm(summary_stat$vapr ~ summary_stat$Marsh_corr_coef)
fit_ph = lm(summary_stat$ph ~ summary_stat$Marsh_corr_coef)
fit_n = lm(summary_stat$n ~ summary_stat$Marsh_corr_coef)
fit_soc = lm(summary_stat$soc ~ summary_stat$Marsh_corr_coef)
fit_sand = lm(summary_stat$sand ~ summary_stat$Marsh_corr_coef)


# For correlation Coefficient
for (i in 8:20){
  # fit a line
  summary_stat$temp = summary_stat[[i]]
  fit = lm(summary_stat$temp ~ summary_stat$Marsh_corr_coef)
  
  # graph

  print(ggplot(summary_stat, aes(x = temp, y = Marsh_corr_coef)) + geom_point() + geom_smooth(method = "lm", se = FALSE) + xlab(colnames(summary_stat[(i)])))
}


# For RMS error
for (i in 8:20){
  # fit a line
  summary_stat$temp = summary_stat[[i]]
  fit = lm(summary_stat$temp ~ summary_stat$Marsh_rms_error)
  
  # graph
  print(ggplot(summary_stat, aes(x = temp, y = Marsh_rms_error)) + geom_point() + geom_smooth(method = "lm", se = FALSE) + xlab(colnames(summary_stat[(i)])))

}

```

-- Not really using anymore -- 

Plot each one over time
```{r}

for (i in 1:nrow(matched)){
  # make a graph of a year of one of them
  temp_data <- filter(daily_sr, Site_ID == matched$site_id[i])
  
  print(ggplot(temp_data) +  
    geom_line(aes(x = Time, y =Marshall_SR, group = 1), linewidth = 1, color = "darkgreen") + 
    geom_line(aes(x = Time, y = Forecast_SR,group = 1), linewidth = 1, color = "orange") +
    geom_line(aes(x = Time, y =  Mill_Quirk_SR, group = 1), linewidth = 1, color = "lightgreen") +
    labs(title = matched$site_id[i], x = "date", y = "Soil Respiration (umol/ (m^2 s)"))
}
```


Compare Cross Correlation Coefficients of sites that lag
```{r}
lag <- data.frame(lag_sites = character(8),
                 half_lag = numeric(8),
                 year_lag = numeric(8))

lag$lag_sites <- c("MLBS", "HARV","SERC","UNDE","ABBY", "GRSM","WREF","SJER")

peak <- numeric(length(lag$lag_sites))
count = 1

for (site in lag$lag_sites){
  temp_data <- filter(daily_sr, Site_ID == site)
  
   # remove na values
  temp_data <- na.omit(temp_data)
  
  name <- temp_data$Site_ID[1]
  # get rid of first row (so i can remove infinite)
  temp_data <- subset(temp_data, select = -c(Site_ID, Time))
  
  # remove inf values
  temp_data <- temp_data[apply(temp_data, 1, function(row) all(is.finite(row))), ]
  
  ccf_data <- ccf(temp_data$Marshall_SR, temp_data$Forecast_SR, lag = 183, plot = TRUE, ylim = range(-1,1))
  ccf_data_2 <- ccf(temp_data$Marshall_SR, temp_data$Forecast_SR, lag = 365, plot = TRUE, ylim = range(-1,1))
  
  #title(name)
  
  lag$half_lag[count] <- ccf_data$lag[which.max(ccf_data$acf)]
  lag$year_lag[count] <- ccf_data_2$lag[which.max(ccf_data_2$acf)]
  count = count + 1
}

# add titles

# analyze more

```
```{r}
lag_all <- data.frame(lag_sites = character(47),
                 any_lag = numeric(47))

lag_all$lag_sites = matched$site_id
count = 1

for (site in lag_all$lag_sites){
  temp_data <- filter(daily_sr, Site_ID == site)
  
   # remove na values
  temp_data <- na.omit(temp_data)
  
  # get rid of first row (so i can remove infinite)
  temp_data <- subset(temp_data, select = -c(Site_ID, Time))
  
  # remove inf values
  temp_data <- temp_data[apply(temp_data, 1, function(row) all(is.finite(row))), ]
  
  if (nrow(temp_data) != 0){
   # ccf_data <- ccf(temp_data$Marshall_SR, temp_data$Forecast_SR, lag = 91, plot = TRUE, ylim = range(-1,1))
    ccf_data_2 <- ccf(temp_data$Marshall_SR, temp_data$Forecast_SR, lag = 365, plot = TRUE, ylim = range(-1,1))
    
    #title(name)
    
    #lag$season_lag[count] <- ccf_data$lag[which.max(ccf_data$acf)]
    lag_all$any_lag[count] <- ccf_data_2$lag[which.max(ccf_data_2$acf)]
  }
  count = count + 1
}
```


