---
title: "Running Ecosystem Simulations Using PEcAn"
author: "Aritra Dey"
format:
  pdf:
    toc: true
    number-sections: false
    fig-width: 10
    fig-height: 6
    fig-dpi: 300
---

# 1. Introduction {#introduction}

Welcome to this PEcAn workflow notebook! This notebook will guide you through running an ecosystem model using PEcAn's programmatic interface.

## 1.1 What Is PEcAn?

PEcAn (Predictive Ecosystem Analyzer) is a scientific workflow system designed to make ecosystem modeling more transparent, repeatable, and accessible. It helps researchers:

- Run ecosystem models with standardized inputs and outputs
- Perform uncertainty analysis on model parameters
- Compare model predictions with observations
- Share and reproduce scientific workflows

## 1.2 What This Notebook Does

This notebook demonstrates how to:

1. Set up and configure a PEcAn workflow
2. Run an ecosystem model simulation
3. Analyze and visualize the results

### The Scenario Being Modeled:

We are modeling carbon and productivity dynamics at the Niwot Ridge Forest AmeriFlux site ([US-NR1](https://ameriflux.lbl.gov/sites/siteinfo/US-NR1), a high-elevation temperate coniferous forest in Colorado. The model configuration uses the SIPNET process-based ecosystem model, parameterized with a temperate coniferous plant functional type (PFT).

The simulation is run for the full year 2004 (January 1 – December 31) using AmeriFlux LBL meteorological drivers from the Niwot Ridge site. The ensemble setup specifies one model run focusing on net primary productivity (NPP) as the target output variable.

This scenario is designed to be a minimal, reproducible example to demonstrate how to run SIPNET within the PEcAn workflow. In later steps, this same framework can be extended to include more ensemble members, additional PFTs, longer time periods, or alternative meteorological inputs.

This run is based on a study by [Moore et. al. (2007)](https://doi.org/10.1016/j.agrformet.2008.04.013) that uses SIPNET to understand the relationship between water and carbon balance at this site. 



## 1.3 Prerequisites

Before running this notebook, make sure you have:

- All the PEcAn packages installed. You can install all PEcAn packages and their dependencies by running the following command in the root of your PEcAn repository:

```

# Enable repository from pecanproject
options(repos = c(
  pecanproject = 'https://pecanproject.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
# Download and install PEcAn.all in R
install.packages('PEcAn.all')

```

- A valid `pecan.xml` configuration file or use the example provided: `pecan/documentation/tutorials/Demo_1_Basic_Run/pecan.xml`

## 1.4 How to Use This Notebook

1. Each section is clearly marked with a heading
2. Code chunks are provided with explanations
3. You can run the code chunks sequentially
4. Once you have successfully run the demo, you can modify parameters to configure new runs and analyses

**Objective:**

This demo illustrates how to run a basic PEcAn workflow using an R-based Quarto notebook. It will cover loading settings, writing model configuration files, and running model simulations. This approach provides a programmatic alternative to the web-based PEcAn interface for executing ecosystem models.

# 2. Session Info

This section prints your R session information for reproducibility. Having this information at the beginning helps with debugging even if the notebook encounters errors later.

```{r version-info}
PEcAn.all::pecan_version()
```

# 3. Install SIPNET v1.3.0

If you haven't already installed the SIPNET binary, you can do so by running the following code. This will download the SIPNET binary to `demo_outdir/sipnet` and make it executable.

> Note: The `demo_outdir` directory will be created in the root of your PEcAn installation (i.e., at `pecan/demo_outdir/`). This directory will contain the SIPNET binary as well as the output generated by PEcAn in this demo.

```{r download-sipnet}
# Download and install SIPNET v1.3.0
source(
  here::here(
    "documentation/tutorials/Demo_1_Basic_Run/download_sipnet.R"
  )
)
```

> Note: You can find the most recent version of the SIPNET binary at: [SIPNET GitHub Releases](https://github.com/PecanProject/sipnet/releases), but this notebook is designed to work with SIPNET v1.3.0.

# 4. Load PEcAn Packages

First, we need to load the PEcAn R packages. These packages provide all the functions we'll use to run the workflow.

```{r libraries}
# Load the PEcAn.all package, which includes all necessary PEcAn functionality
library("PEcAn.all")
```

# 5. Load PEcAn Settings File

PEcAn uses an XML-based settings file (pecan.xml) to configure model runs. This file defines key information about the run including: PFT(s), site location, time period of the run, the location of input files and where outputs will be saved. Other settings outside the scope of this demo include the types of analyses that will be performed, how to connect to a database, and how to run it on a high performance computing cluster  (we are using the default single model run on a single computer). 

You can read more about the settings file in the ["PEcAn XML" chapter](https://pecanproject.github.io/pecan-documentation/develop/pecanXML.html#pecanXML) of the documentation.

There is an example `pecan.xml` that has been configured for this demonstration. You can find it at `pecan/documentation/tutorials/Demo_1_Basic_Run/pecan.xml`.


This is how PEcAn loads the settings file:

```{r load-settings}
settings_path <- here::here("documentation/tutorials/Demo_1_Basic_Run/pecan.xml")
```

# 6. Prepare and Validate Settings

After specifying the path to the `pecan.xml` file, the next step involves reading and preparing these settings. PEcAn provides utilities to process and validate the configurations before execution begins. 

- `PEcAn.settings::read.settings(settings_path)`: Reads the `pecan.xml` file and converts it to an R list object.
- `PEcAn.settings::prepare.settings(settings)`: Prepares and validates settings. It sets defaults for missing fields, changes file paths to absolute paths, and generally ensures consistency.

```{r read-prepare-settings}
# Read the settings from the pecan.xml file
settings <- PEcAn.settings::read.settings(settings_path)

# Prepare and validate the settings
settings <- PEcAn.settings::prepare.settings(settings)
```

# 7. Explore the Settings Object

Once the settings have been read and prepared, it is useful to inspect the structure of the `settings` object. This object is an R list containing all parameters and configurations for the PEcAn workflow.

*   `str(settings)` displays the internal structure of the `settings` object. This shows how the settings are represented in R and is useful for debugging and verifying settings.

```{r explore-settings}
str(settings)
```

Your turn: explore the `settings` object further using the following commands:
*   `names(settings)` to list the top-level keys in the settings object
*   Once you know the names, you can look at each component in detail, for example:
*   `settings$run` to access the run-specific settings, such as start and end dates, model type, and output directory
*   `settings$pfts` to explore the Plant Functional Types settings

Now you can update each of these settings. Here is a simple example:

```{r add-info-to-settings}
settings$info <- list(
  author = "Aritra Dey",
  date = Sys.Date(),
  description = "Demo run of PEcAn using SIPNET"
)
```

Editing the more interesting settings to change the PFT (`settings$pfts`) or extend the run (`settings$run$end.date`) is beyond the scope of this demo. You _could_ change the pft or the end date, but you would need a new file containing parameters for that PFT (`settings$pfts$pft$posterior.files`), or a climate file (`settings$run$met$path$path1`) that extends to the desired simulation period.

# 8. Create Output Directory

Before running the workflow, we need to ensure that the output directory specified in the settings exists. This directory will store all the model outputs, configuration files, and results from the simulation.

> **Note:** All outputs will be written to `settings$outdir`. You can change this directory in your `pecan.xml` file or by modifying `settings$outdir` in the notebook before running the workflow. You can also put the `sipnet` executable anywhere you prefer, as long as you update `settings$model$binary`.

```{r create-outdir}
dir.create(settings$outdir, recursive = TRUE, showWarnings = FALSE)
```

# 9. Write Model Configuration Files

This step generates the model-specific configuration files that will be used to run the ecosystem model. The process involves:

1. Disabling database write operations because we are not using a database
2. Generating SIPNET configuration files using the `runModule.run.write.configs()` function.

```{r write-configs}
settings$database <- NULL # Disable database operations for this demo
settings <- PEcAn.workflow::runModule.run.write.configs(settings)
```

# 10. Run Model Simulations and Fetch Results

This section executes the actual model simulations and retrieves the results. The process is managed by PEcAn's workflow system, which handles the execution of your chosen ecosystem model.

*   `runModule_start_model_runs(settings)`: This function initiates the model runs based on your configuration. It manages the execution of your chosen ecosystem model, using the configuration files generated in the previous step.

```{r run-model}
PEcAn.workflow::runModule_start_model_runs(settings)
```

# 11. Extract Model Results and Prepare for Analysis

After the model simulation completes, we need to extract the results and prepare them for analysis. This involves:

1. Reading the run ID
2. Setting up output paths
3. Defining time period
4. Loading model output

```{r get-plot-vars}
runid <- as.character(read.table(paste(settings$outdir, "/run/", "runs.txt", sep = ""))[1, 1]) # Note: if you are using an xml from a run with multiple ensembles this line will provide only the first run id
outdir <- file.path(settings$outdir, "/out/", runid)
start.year <- lubridate::year(settings$run$start.date)
end.year <- lubridate::year(settings$run$end.date)
model_output <- PEcAn.utils::read.output(
  runid,
  outdir,
  start.year,
  end.year,
  variables = NULL,
  dataframe = TRUE,
  verbose = FALSE
)
available_vars <- names(model_output)[!names(model_output) %in% c("posix", "time_bounds")]
```

# 12. Display Available Model Variables

This section shows all the variables that are available in the model output. These variables represent different ecosystem processes and states that the model has simulated.

```{r show-variables, echo=FALSE}
vars_df <- PEcAn.utils::standard_vars |>
  dplyr::select(
    Variable = Variable.Name,
    Description = Long.name
  ) |>
  dplyr::filter(Variable %in% available_vars) |>
  # TODO: add year to PEcAn.utils::standard vars
  dplyr::bind_rows(
    dplyr::tibble(
      Variable = "year",
      Description = "Year"
    )
  )

vars_df$Description[is.na(vars_df$Description)] <- "(No description available)"
knitr::kable(vars_df, caption = "Model Output Variables and Descriptions")
```

# 13. Visualize Model Results

This section provides examples of how to create time series plots of different model variables. The examples cover various ecosystem processes including carbon fluxes, carbon pools, water variables, and structural variables like Leaf Area Index (LAI).

## 13.1 Plot Carbon Fluxes

```{r plot-carbon-fluxes}
# Plot Gross Primary Productivity (GPP) and Net Primary Productivity (NPP)
plot(model_output$posix, model_output$GPP,
  type = "l",
  col = "green",
  xlab = "Date",
  ylab = "Carbon Flux (kg C m-2 s-1)",
  main = "Carbon Fluxes Over Time"
)
lines(model_output$posix, model_output$NPP, col = "blue")
legend("topright", legend = c("GPP", "NPP"), col = c("green", "blue"), lty = 1)
```

## 13.2 Plot Carbon Pools

```{r plot-carbon-pools}
# Plot Total Live Biomass and Total Soil Carbon
plot(model_output$posix, model_output$TotLivBiom,
  type = "l",
  col = "darkgreen",
  xlab = "Date",
  ylab = "Carbon Pool (kg C m-2)",
  main = "Carbon Pools Over Time"
)
lines(model_output$posix, model_output$TotSoilCarb, col = "brown")
legend("topright", legend = c("Total Live Biomass", "Total Soil Carbon"), col = c("darkgreen", "brown"), lty = 1)
```

## 13.3 Plot Water Variables

```{r plot-water-variables}
# Plot Soil Moisture and Snow Water Equivalent
plot(model_output$posix, model_output$SoilMoist,
  type = "l",
  col = "blue",
  xlab = "Date",
  ylab = "Soil Moisture (kg m-2)",
  main = "Soil Moisture Over Time"
)
lines(model_output$posix, model_output$SWE, col = "lightblue")
legend("topright", legend = c("Soil Moisture", "Snow Water Equivalent"), col = c("blue", "lightblue"), lty = 1)
```

## 13.4 Plot LAI and Biomass

```{r plot-lai-biomass}
# Plot Leaf Area Index (LAI) and Above Ground Wood
plot(model_output$posix, model_output$LAI,
  type = "l",
  col = "darkgreen",
  xlab = "Date",
  ylab = "LAI (m2 m-2)",
  main = "Leaf Area Index Over Time"
)
lines(model_output$posix, model_output$AbvGrndWood, col = "brown")
legend("topright", legend = c("LAI", "Above Ground Wood"), col = c("darkgreen", "brown"), lty = 1)
```

# 14. PEcAn Outputs

A number of files generated by the underlying ecosystem model are available for exploration and further use. These include:

1. Output files in the standardized netCDF ([year].nc) that can be downloaded for visualization and analysis (R, Matlab, ncview, panoply, etc)
2. Raw model output in model-specific format (e.g. sipnet.out).
3. Logfile.txt contains job.sh & model error, warning, and informational messages

## Directory Structure

These are the key folders and files that will be created under the directory defined by `settings$outdir` (e.g., `demo_outdir` in the example). The file contents are described in the next section.

```
demo_outdir/
├── run/                   # Configuration & execution metadata
│   ├── runs.txt           # List of run IDs (one per model realization)
│   ├── <runid>/           # Model-specific config copies (sometimes)
│   └── config.*           # Generated model configs (e.g., SIPNET)
├── out/                   # Raw model outputs by run ID
│   └── <runid>/           # E.g., daily or sub-daily SIPNET output files
```

## Output Files

The following sections describe the contents of these folders and files in more detail.

### Traits 

**The `/pft` directory** contains an array of outputs available that are related to the process of estimating the model parameters. 

The Rdata files **trait.data.Rdata** and **jagged.data.Rdata** are, respectively, the available trait data extracted from the database that was used to estimate the model parameters and that same data cleaned and formatted for the statistical code. The **list of variables that are queried is determined by what variables have priors associated with them in the definition of the PFTs**. Priors are output into **prior.distns.Rdata**. Likewise, the **list of species that are associated with a PFT determines what subset of data is extracted** out of all data matching a given variable name. Demo 3 will demonstrate how a PFT can be created or modified. To look at these files in RStudio **click on these files to load them into your workspace**. You can further examine them in the _Environment_ window or accessing them at the command line. For example, try typing ```names(trait.data)``` as this will tell you what variables were extracted, ```names(trait.data$Amax)``` will tell you the names of the columns in the Amax table, and ```summary(trait.data$Amax)``` will give you summary data about the Amax values.


### META-ANALYSIS:

- `*.bug`: The evaluation of the meta-analysis is done using a Bayesian statistical software package called JAGS that is called by the R code. For each trait, the R code will generate a [trait].model.bug file that is the JAGS code for the meta-analysis itself. This code is generated on the fly, with PEcAn adding or subtracting the site, treatment, and greenhouse terms depending upon the presence of these effects in the data itself. If the <random.effects> tag is set to FALSE then all random effects will be turned off even if there are multiple sites. 
- `meta-analysis.log` contains a number of diagnostics, including the summary statistics of the model, an assessment of whether the posterior is consistent with the prior, and the status of the Gelman-Brooks-Rubin convergence statistic (which is ideally 1.0 but should be less than 1.1). 
- `ma.summaryplots.*.pdf` are collections of diagnostic plots produced in R after the above JAGS code is run that are useful in assessing the statistical model. _Open up one of these pdfs_ to evaluate the shape of the posterior distributions (they should generally be unimodal), the convergence of the MCMC chains (all chains should be mixing well from the same distribution), and the autocorrelation of the samples (should be low).
- `traits.mcmc.Rdata` contains the raw output from the statistical code. This includes samples from all of the parameters in the meta-analysis model, not just those that feed forward to the ecosystem, but also the variances, fixed effects, and random effects. 
- `post.distns.Rdata` stores a simple table of the posterior distributions for all model parameters in terms of the name of the distribution and its parameters. 
- `posteriors.pdf` provides graphics showing, for each model parameter, the prior distribution, the data, the smoothed histogram of the posterior distribution (labeled post), and the best-fit analytical approximation to that smoothed histogram (labeled approx). _Open posteriors.pdf and compare the posteriors to the priors and data_.


### Additional Workflow Outputs:

The root workflow folder contains logs and records files from the PEcAn workflow. They provide a detailed record of how data was generated and are key components of the analysis metadata & provenance. These can be useful for debugging errors as well as for downstream analysis.

In addition to log files, key outputs include:

1. `STATUS` a summary of the steps of the workflow, the time they took, and whether they were successful
2. `pecan.*.xml` a copy of the prepared PEcAn settings file.

# 15. Conclusion

This notebook demonstrated how to set up, run, and analyze a PEcAn ecosystem model workflow programmatically. You can now modify parameters, try different models, or extend the analysis as needed.

Try editing the `pecan.xml` file. Give it a new name and update the **settings_path** variable at the beginning of this Demo to point to the new file. See how the changes affect the model output!

# 15. Clean Up Workflow Output (Optional)

If you want to remove all files and directories created by this workflow and start fresh, you can run the following code. This will delete the entire output directory specified in your settings. **Use with caution!**

```{r cleanup}
# WARNING: This will permanently delete all workflow output files!
# Uncomment the line below to enable cleanup.
# fs::dir_delete(settings$outdir)
```

# 16. Session Info
This section prints your R session information for reproducibility.

```{r session-info}
sessionInfo()
```

